---
title: "ClassIII_Peptides"
author: "Tülay Karakulak"
date: "2025-02-25"
output: html_document
---

Ovarian Samples ClassIII Peptides
Criteria:
- FDR < 0.1. FDR correction is per gene.
- Peptides either coming from cMDT or MDT or from both are counted, they should show a differential expression.
- cMDT genes that are not found in 1st and 2nd Classes.

#required libraries
```{r loadlibraries, echo=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(sva)
library(stats)
library(data.table)
library(BSDA)
library(Biostrings)
library(ggvenn)
```

We decided to use the normalized value instead of the quantity - and divided by Isoform number.

```{r classIIIpeptideCalc2}
ClassIII_calc_Norm2 <- function(Main_output_filtered_peptides, DT, Switches_df, all_samples_isolate) {
  
results <- data.frame(ENSG = as.character(), MeanCancer = as.numeric(), MeanNormal = as.numeric(), p_value = as.numeric())

for (eachPeptide in unique(Main_output_filtered_peptides$Stripped.Sequence)) {
  
  ENSG_id  <- DT[DT$Stripped.Sequence == eachPeptide, 'Genes']$Genes[1]
  # we should only look at the common samples - and isolate those

  Samples_w_ensg <- Switches_df[Switches_df$ENSG == ENSG_id, 'SampleID']
  sample_ids_found <- sapply(strsplit(Samples_w_ensg, "/"), "[", 2)
  samples_ids_nonfound <- setdiff(all_samples_isolate, sample_ids_found)
  
  sample_ids_found_quantity <- DT[DT$Sampleid %in% sample_ids_found & DT$Stripped.Sequence ==  eachPeptide, 'NormalizedByGene']
  
  if (length(sample_ids_found_quantity$NormalizedByGene) > 1 ) {
    
    Isoform <- DT[DT$Sampleid %in% sample_ids_found & DT$Stripped.Sequence ==  eachPeptide, 'Protein.Ids']$Protein.Ids
    NumberofIsoforms <- length(strsplit(Isoform, ';')[[1]])
  
    sample_ids_found_quantity2 <- (sample_ids_found_quantity$NormalizedByGene)/NumberofIsoforms
    
    sample_ids_found_quantity_mean <- mean(sample_ids_found_quantity2)
  } else if (length(sample_ids_found_quantity$NormalizedByGene) == 1) {
    sample_ids_found_quantity_mean <- 'OneSample'
  } else {
    sample_ids_found_quantity_mean <- NA
  }
  
  sample_ids_nonfound_quantity <- DT[DT$Sampleid %in% samples_ids_nonfound & DT$Stripped.Sequence ==  eachPeptide, 'NormalizedByGene',]
  
  if(length(sample_ids_nonfound_quantity$NormalizedByGene) > 1 ) {
    
    Isoform <- DT[DT$Sampleid %in% samples_ids_nonfound & DT$Stripped.Sequence ==  eachPeptide, 'Protein.Ids']$Protein.Ids
    
    NumberofIsoforms <- length(strsplit(Isoform, ';')[[1]])
  
    sample_ids_nonfound_quantity2 <- (sample_ids_nonfound_quantity$NormalizedByGene)/NumberofIsoforms
    
    sample_ids_nonfound_quantity_mean <- mean(sample_ids_nonfound_quantity2)
    
  } else if (length(sample_ids_nonfound_quantity$NormalizedByGene) == 1){
    sample_ids_nonfound_quantity_mean <- 'OneSample'
  } else {
    sample_ids_nonfound_quantity_mean <- NA
  }

  if(length(sample_ids_found_quantity$NormalizedByGene) > 1 & length(sample_ids_nonfound_quantity$NormalizedByGene) > 1) {
   testing <- wilcox.test(as.numeric(unlist(sample_ids_found_quantity2)), as.numeric(unlist(sample_ids_nonfound_quantity2)), alternative = "two.sided", paired = FALSE, exact = NULL, correct = TRUE)
   p_value <- testing$p.value
  } else {
    p_value <- NA
    
  }
  
  results_eachPeptide <- data.frame(ENSG = ENSG_id, Peptide = eachPeptide, MeanCancer = sample_ids_found_quantity_mean, MeanNormal = sample_ids_nonfound_quantity_mean, p_value = p_value)

  results <- rbind(results, results_eachPeptide)  

} 
  return(results)
  
  }
```



```{r merge_dMDTs_ENSPs}
ensg_enst_ensp_apris_v104 <-  read.csv('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Biomart/mart_export-2May2023.txt', sep='\t')
head(ensg_enst_ensp_apris_v104)
```


## Ovarian Samples
```{r read_output_diann_Ovary}
Main_output_Ovary <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/Ovary/DIANN/Main_Report.tsv', delim = '\t', col_names = TRUE)

Main_output_Ovary <- Main_output_Ovary %>% filter(!str_detect(Genes, ";"))

Main_output_filtered_Ovary <- Main_output_Ovary[Main_output_Ovary$Q.Value <= 0.01 & Main_output_Ovary$PG.Q.Value <= 0.01, ]

head(Main_output_filtered_Ovary)
dim(Main_output_filtered_Ovary)
```


```{r read_MDTs}
Switches_Ovary_M <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Met_GTEx.RDS')
Switches_Ovary_P <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Pri_GTEx.RDS')
Switches_Ovary <- rbind(Switches_Ovary_P, Switches_Ovary_M)
```


```{r merge_ensts}
Switches_Ovary_ensps <- merge(ensg_enst_ensp_apris_v104, Switches_Ovary, by.x= 'Transcript.stable.ID', by.y = 'dMDT', sort=FALSE)
head(Switches_Ovary_ensps)
```


```{r mdt_enst_merging}
Switches_Ovary_MDTS_splitted <- Switches_Ovary %>%
  mutate(MDT_GTEx = strsplit(as.character(MDT_GTEx), ",")) %>%
  unnest(MDT_GTEx)

Switches_Ovary_ensps_mdts <- merge(ensg_enst_ensp_apris_v104, Switches_Ovary_MDTS_splitted, by.x= 'Transcript.stable.ID', by.y = 'MDT_GTEx', sort=FALSE)
head(Switches_Ovary_ensps_mdts)
```


```{r check_dMDT_GTEx}
# filter main output
Main_output_filtered_dMDT <- Main_output_filtered_Ovary[Main_output_filtered_Ovary$Protein.Ids %in% Switches_Ovary_ensps$Protein.stable.ID, ]

head(Main_output_filtered_dMDT)
dim(Main_output_filtered_dMDT)
#[1] 7561   58

unique_peptides <- Main_output_filtered_dMDT[,c('Protein.Ids','Stripped.Sequence')] %>% dplyr::distinct()
```


```{r upper_code_vectorizaition}
library(data.table)
dMDT_vector <- unique(Switches_Ovary_ensps$Protein.stable.ID)
MDT_vector <- unique(Switches_Ovary_ensps_mdts$Protein.stable.ID)

dMDT_proteomics <- Main_output_filtered_Ovary[FALSE, ]
MDT_proteomics <- Main_output_filtered_Ovary[FALSE, ]

# Convert data frames to data.table
Main_output_filtered_dt <- as.data.table(Main_output_filtered_Ovary)
dMDT_proteomics_dt <- as.data.table(dMDT_proteomics)
MDT_proteomics_dt <- as.data.table(MDT_proteomics)

# Create logical vectors indicating matching IDs
dMDT_match <- sapply(Main_output_filtered_dt$Protein.Ids, function(ids) any(str_split_1(ids, ';') %in% dMDT_vector)) # Protein IDs that include dMDT
MDT_match <- sapply(Main_output_filtered_dt$Protein.Ids, function(ids) any(str_split_1(ids, ';') %in% MDT_vector)) # Protein IDs that include MDT

# Filter rows based on the logical vectors
dMDT_proteomics_dt <- Main_output_filtered_dt[dMDT_match]
MDT_proteomics_dt <- Main_output_filtered_dt[MDT_match]
```


```{r check_dMDTs_in_general}
library(dplyr)

# check rows dMDT is found together with other ENSP ids but excluding the MDTs 
library(purrr)
include_ids <- unique(Switches_Ovary_ensps$Protein.stable.ID)
exclude_ids <- unique(Switches_Ovary_ensps_mdts$Protein.stable.ID)

# Peptides that belong to dMDT and others but not to MDT ids
Main_output_filtered_dMDT_exc_MDT <- Main_output_filtered_Ovary %>%
  dplyr::filter(
    str_detect(Protein.Ids, paste(include_ids, collapse = '|')) &
    !str_detect(Protein.Ids, paste(exclude_ids, collapse = '|'))
  )

dim(Main_output_filtered_dMDT_exc_MDT)
#[1] 38103    58
length(unique(Main_output_filtered_dMDT_exc_MDT$Protein.Ids))
#[1] 361 # unique dMDT number found in proteomics data is 121 but with other transcripts exlucidng the MDT is 361 
```


```{r CheckPeptideNumberPerGene}

# Calculate number of peptides per gene
Main_output_filtered_peptides_Ovary <- Main_output_filtered_Ovary %>% dplyr::select(Run, Stripped.Sequence, Genes) %>% dplyr::filter(Genes %in% Switches_Ovary$ENSG) %>% dplyr::distinct() %>% dplyr::group_by(Stripped.Sequence, Genes) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count)) 

DT_Ovary <- as.data.table(Main_output_filtered_Ovary)
DT_Ovary[, Sampleid := sapply(DT_Ovary$File.Name, function(x) strsplit(strsplit(x, "/")[[1]][11], '-')[[1]][1])]

all_samples_ovary <- unique(Switches_Ovary$SampleID)
all_samples_isolate_ovary <- sapply(strsplit(all_samples_ovary, "/"), "[", 2)
```


```{r NormalizeAllPeptideExpbyGene}
DT_Ovary_SumStrippedSequence <- DT_Ovary %>% dplyr::select(Sampleid, Protein.Ids, Genes, Precursor.Normalised, Stripped.Sequence) %>% dplyr::group_by(Sampleid, Protein.Ids, Genes, Stripped.Sequence) %>% dplyr::mutate(AvgNormalised = mean(Precursor.Normalised)) %>% dplyr::ungroup() %>% dplyr::select(Sampleid, Protein.Ids, Genes, Stripped.Sequence, AvgNormalised) %>% dplyr::distinct()

DT_Normalized_Ovary <- DT_Ovary_SumStrippedSequence %>% dplyr::select(Protein.Ids, Genes, AvgNormalised, Stripped.Sequence, Sampleid) %>% dplyr::distinct() %>% dplyr::group_by(Sampleid, Genes) %>% dplyr::mutate(NormalizedByGene = AvgNormalised/sum(AvgNormalised))
```


```{r Ovary_ClassIII}
ClassIII_Ovary2 <- ClassIII_calc_Norm2(Main_output_filtered_peptides_Ovary, DT_Normalized_Ovary, Switches_Ovary, all_samples_isolate_ovary)
```

```{r pval}
ClassIII_Ovary_fdr <- ClassIII_Ovary2 %>% dplyr::group_by(ENSG) %>% dplyr::mutate(FDR_new = p.adjust(p_value, method = "BH"))
```

```{r mergefiles}
ClassIII_Ovary_fdr_only_proteinIds <- (merge(ClassIII_Ovary_fdr, Main_output_filtered_Ovary[, c('Protein.Ids', 'Genes', 'Stripped.Sequence')], by.x = 'Peptide', by.y = 'Stripped.Sequence') %>% dplyr::distinct())
head(ClassIII_Ovary_fdr_only_proteinIds)
```


```{r CategoriseENSPs}

for (eachrow in 1:nrow(ClassIII_Ovary_fdr_only_proteinIds)) { 
  protein_ids <- ClassIII_Ovary_fdr_only_proteinIds[eachrow, 'Protein.Ids']
  protein_ids_vector <- unlist(str_split(protein_ids, ';'))
  
  # Check if any protein ID is in exclude_ids or include_ids
  in_MDT <- any(protein_ids_vector %in% exclude_ids)
  in_cMDT <- any(protein_ids_vector %in% include_ids)
  
  if (in_MDT & !in_cMDT) {
    ClassIII_Ovary_fdr_only_proteinIds$description[eachrow] <- "MDT, not cMDT"
  } else if (!in_MDT & in_cMDT) {
    ClassIII_Ovary_fdr_only_proteinIds$description[eachrow] <- "cMDT, not MDT" # This should not be found considering that they are Class II peptides - as a control!
  } else if (in_MDT & in_cMDT) {
    ClassIII_Ovary_fdr_only_proteinIds$description[eachrow] <- "MDT and cMDT"
  } else {
    ClassIII_Ovary_fdr_only_proteinIds$description[eachrow] <- "Other isoforms"
  }
}
```


```{r saveData}
saveRDS(ClassIII_Ovary_fdr_only_proteinIds, '/Users/admin/Desktop/TuPro_Paper/BJC/Revision/ClassIII_Ovary_fdr_only_proteinIds_StrippedSequence_NormalizedByPeptideNumberAndRelativeExp.RDS')
```

ClassIII_Ovary_fdr_only_proteinIds %>% dplyr::filter(Protein.Ids %in% Switches_Ovary_ensps$Protein.stable.ID) %>% dplyr::select(Protein.Ids) %>% dplyr::distinct() # 121 Class I peptides

ClassIII_Ovary_fdr_only_proteinIds[ClassIII_Ovary_fdr_only_proteinIds$description == 'cMDT, not MDT', ] %>% separate_rows(Protein.Ids, sep = ";") %>% dplyr::filter(Protein.Ids %in% Switches_Ovary_ensps$Protein.stable.ID) %>% dplyr::select(Protein.Ids) %>% dplyr::distinct()

## Ovarian Samples - Class I and Class II peptides list
```{r ClassIandII}
# Main_output_filtered_dMDT_exc_MDT # it is from Manuscript4.Rmd file
# ENSGs with only 1 cMDTs and the ones that are found in Class I or Class II peptides list. We should remove these ENSGs from the list
ENSG_list_to_remove <- Switches_Ovary %>% dplyr::select(ENSG, dMDT) %>% dplyr::distinct() %>% dplyr::group_by(ENSG) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count)) %>% dplyr::filter(count == 1) %>% dplyr::filter(ENSG %in% Main_output_filtered_dMDT_exc_MDT$Genes)

ClassIII_Ovary_fdr_only_ClassIII <- ClassIII_Ovary_fdr_only_proteinIds %>% dplyr::filter(!Peptide %in% Main_output_filtered_dMDT_exc_MDT$Stripped.Sequence) %>% dplyr::filter(!ENSG %in% ENSG_list_to_remove$ENSG)# Peptides that are not found in Class I and Class II! 

head(ClassIII_Ovary_fdr_only_ClassIII)
```


```{r saveData}
saveRDS(ClassIII_Ovary_fdr_only_ClassIII, '/Users/admin/Desktop/TuPro_Paper/BJC/Revision/ClassIII_Ovary_Peptides.RDS')
```
dim(ClassIII_Ovary_fdr_only_ClassIII)
[1] 18035     9


## 4 category search
```{r Categories1_3}
# MDT specific peptide doesn’t change, cMDT specific changes
# This category is looking at the isoforms that have both 'MDT, not cMDT' and 'cMDT and MDT' categories
valid_ENSGs <- ClassIII_Ovary_fdr_only_ClassIII %>% dplyr::select(Peptide, ENSG, MeanCancer, MeanNormal, p_value, FDR_new, Protein.Ids, description) %>% dplyr::group_by(ENSG) %>% dplyr::summarise(has_both = all(c('MDT and cMDT', 'MDT, not cMDT') %in% description)) %>% dplyr::filter(has_both) %>% dplyr::pull(ENSG)

Category1_3_Categories <-  ClassIII_Ovary_fdr_only_ClassIII %>%
  filter(ENSG %in% valid_ENSGs) %>% dplyr::filter(description %in% c('MDT and cMDT', 'MDT, not cMDT'))
```

```{r Category1_4}
# Define FDR significance threshold
fdr_threshold <- 0.1

# Identify ENSG values where 'MDT, not cMDT' has significant FDR and 'MDT and cMDT' has no change
# category 2
genes_with_significant_change <- Category1_3_Categories %>%
  group_by(ENSG) %>%
  summarise(
    has_significant_not_cMDT = any(FDR_new[description == "MDT, not cMDT"] <= fdr_threshold, na.rm = TRUE),
    has_no_significant_cMDT = all(FDR_new[description == "MDT and cMDT"] %>% na.omit() > fdr_threshold)
  ) %>%
  filter(has_significant_not_cMDT & has_no_significant_cMDT) %>%
  pull(ENSG)

# category 1
genes_with_significant_change_ctg1 <- Category1_3_Categories %>%
  group_by(ENSG) %>%
  summarise(
    has_significant_not_cMDT = any(FDR_new[description == "MDT, not cMDT"] > fdr_threshold, na.rm = TRUE),
    has_no_significant_cMDT = all(FDR_new[description == "MDT and cMDT"] %>% na.omit() <= fdr_threshold)
  ) %>%
  filter(has_significant_not_cMDT & has_no_significant_cMDT) %>%
  pull(ENSG)

# category 3
genes_with_significant_change_ctg3 <- Category1_3_Categories %>%
  group_by(ENSG) %>%
  summarise(
    has_significant_not_cMDT = any(FDR_new[description == "MDT, not cMDT"] <= fdr_threshold, na.rm = TRUE),
    has_no_significant_cMDT = all(FDR_new[description == "MDT and cMDT"] %>% na.omit() <= fdr_threshold)
  ) %>%
  filter(has_significant_not_cMDT & has_no_significant_cMDT) %>%
  pull(ENSG)

# Filter the dataframe to retain only these genes
Category1 <- Category1_3_Categories %>%
  filter(ENSG %in% genes_with_significant_change_ctg1)

Category2 <- Category1_3_Categories %>%
  filter(ENSG %in% genes_with_significant_change)

Category3 <- Category1_3_Categories %>%
  filter(ENSG %in% genes_with_significant_change_ctg3)
```

length(unique(Category2$ENSG)) 11
length(unique(Category1$ENSG)) 44
length(unique(Category3$ENSG)) 7 

```{r category4}
# where there is no shared cMDT and MDT peptides. - and only change in the MDT specific peptides.
valid_ENSGs <- ClassIII_Ovary_fdr_only_ClassIII %>% dplyr::select(Peptide, ENSG, MeanCancer, MeanNormal, p_value, FDR_new, Protein.Ids, description) %>% dplyr::group_by(ENSG) %>% dplyr::summarise(has_both = all(c('MDT and cMDT', 'MDT, not cMDT') %in% description)) %>% dplyr::filter(has_both) %>% dplyr::pull(ENSG)
```


```{r Category4}
# FDR < 0.1 or MeanNormal has a value but MeanCancer does not have a value
valid_ENSGs_only_not_cMDT <- ClassIII_Ovary_fdr_only_ClassIII %>%
  dplyr::select(Peptide, ENSG, MeanCancer, MeanNormal, p_value, FDR_new, Protein.Ids, description) %>%
  dplyr::group_by(ENSG) %>%
  dplyr::summarise(
    has_not_cMDT = any(description == 'MDT, not cMDT'),
    has_cMDT = any(description == 'MDT and cMDT'),
    has_only_Normal = any(!is.na(MeanNormal) & is.na(MeanCancer))
  ) %>%
  dplyr::filter(has_not_cMDT & !has_cMDT) %>%
  dplyr::pull(ENSG)

Category4 <-  ClassIII_Ovary_fdr_only_ClassIII %>%
  filter(ENSG %in% valid_ENSGs_only_not_cMDT) %>%
  filter(
    !(MeanNormal == "NA" | MeanNormal == "OneSample") & 
    (is.na(MeanCancer) | MeanCancer != "OneSample") &
    # Remove cases where MeanNormal is "NA"
    !(MeanNormal == "OneSample" & MeanCancer == "NA")  &
      (is.na(FDR_new) | FDR_new <= 0.1) # Remove cases where MeanNormal is "OneSample" and MeanCancer is "NA"
  )
```


```{r merge_categories}
Category1$Category <- 'Category1'
Category2$Category <- 'Category2'
Category3$Category <- 'Category3'
Category4$Category <- 'Category4'
All4Categories <- rbind(Category1, Category2, Category3, Category4)
write.table(All4Categories, '/Users/admin/Desktop/TuPro_Paper/BJC/Revision/ClassIII_FourCategory_Ovary.txt', quote = FALSE)
```


