---
title: "TCGA_Mutation"
author: "Tülay Karakulak"
date: "2024-June-04"
output: html_document
---

# Spliceosome and Mutation Analysis - Supplementary Figure 2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(TCGAbiolinks)
library(dplyr)
library(DT)
library(maftools)
library(ggplot2)
library(hrbrthemes)
```


```{r query}
query <- GDCquery(
    project = "TCGA-SKCM", 
    data.category = "Simple Nucleotide Variation", 
    access = "open",
    data.type = "Masked Somatic Mutation", 
    workflow.type = "Aliquot Ensemble Somatic Variant Merging and Masking"
)
GDCdownload(query)
maf <- GDCprepare(query)

query_Ovary <- GDCquery(
    project = "TCGA-OV", 
    data.category = "Simple Nucleotide Variation", 
    access = "open",
    data.type = "Masked Somatic Mutation", 
    workflow.type = "Aliquot Ensemble Somatic Variant Merging and Masking"
)
GDCdownload(query_Ovary)
maf_Ovary <- GDCprepare(query_Ovary)


query_AML <- GDCquery(
    project = "TCGA-LAML", 
    data.category = "Simple Nucleotide Variation", 
    access = "open",
    data.type = "Masked Somatic Mutation", 
    workflow.type = "Aliquot Ensemble Somatic Variant Merging and Masking"
)
GDCdownload(query_AML)
maf_AML <- GDCprepare(query_AML)
```


```{r db}
#maf <- GDCprepare(query)
#datatable(maf[1:20,],
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#plotmafSummary(maf = maf_AML, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE)
#
#datatable(maf_Ovary[1:20,],
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#
#datatable(maf_AML[1:20,],
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#
#maf <- read.maf(maf = maf)
#datatable(getSampleSummary(maf),
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#plotmafSummary(maf = maf, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE)
#
### Ovary
#maf_Ovary <- read.maf(maf = maf_Ovary)
#
#datatable(getSampleSummary(maf_Ovary),
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#plotmafSummary(maf = maf_Ovary, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE)
#
### AML
#maf_AML <- read.maf(maf = maf_AML)
#datatable(getSampleSummary(maf_AML),
#          filter = 'top',
#          options = list(scrollX = TRUE, keys = TRUE, pageLength = 5), 
#          rownames = FALSE)
#plotmafSummary(maf = maf_AML, rmOutlier = TRUE, addStat = 'median', dashboard = TRUE)
```


```{r maf_df_2}
maf_barcode_ids <- maf %>% dplyr::select(Tumor_Sample_Barcode, Tumor_Sample_UUID) %>% dplyr::distinct()
```


```{r merge_files}
maf <- read.maf(maf = maf)
maf_df <- mafSummary(maf)$variant.classification.summary
maf_barcode_ids_mutations <- merge(maf_barcode_ids, maf_df, by= 'Tumor_Sample_Barcode')
```

```{r read_switches}
dMDT_Melanoma_tcga_primary <- readRDS('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Melanoma/Switches/Switches_TCGA_Pri_GTEx.RDS')

dMDT_Melanoma_tcga_metastatis <- readRDS('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Melanoma/Switches/Switches_TCGA_Met_GTEx.RDS')
```

```{r TCGA_id_convert}
library(TCGAutils)
#gender_info <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/Melanoma/TCGA/clinical.cases_selection.2022-10-07/clinical.tsv', sep = '\t', header=TRUE)
#head(gender_info)

sample_ids <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/Melanoma/TCGA/biospecimen.cases_selection.2022-10-07/aliquot.tsv', sep = '\t', header=TRUE)

#sample_id_case_id <- merge(sample_ids, gender_info, by= 'case_id')

# convert old TCGA ids to new ids
history_updated2 <- data.frame(oldid = character(), newid = character())
for (eachfile in c(unique(dMDT_Melanoma_tcga_primary$SampleID), unique(dMDT_Melanoma_tcga_metastatis$SampleID))) {
  updated_list <- UUIDhistory(eachfile)
  oldfile <- updated_list[1,1]
  newfile <- updated_list[2,1]
  history_updated <- data.frame(oldid = oldfile, newid = newfile)
  history_updated2 <- rbind(history_updated2, history_updated)
}

tcga_case_id <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$newid,to_type = "case_id")
tcga_case_id_rest <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$oldid,to_type = "case_id") ## some ids are still in use in TCGA - not all are updated!
head(tcga_case_id)
```

```{r merge_sample_id}
sample_ids_new <- rbind(sample_ids[sample_ids$case_id %in% tcga_case_id$cases.case_id,], sample_ids[sample_ids$case_id %in% tcga_case_id$file_id,], sample_ids[sample_ids$case_id %in% tcga_case_id_rest$cases.case_id,], sample_ids[sample_ids$case_id %in% tcga_case_id_rest$file_id,])

sample_ids_new_oldIds <- merge(sample_ids_new, tcga_case_id, by.x= "case_id", by.y="cases.case_id")
sample_ids_new_oldIds_2 <- merge(sample_ids_new, tcga_case_id_rest, by.x= "case_id", by.y="cases.case_id")

sample_ids_new_case_ids <- rbind(sample_ids_new_oldIds, sample_ids_new_oldIds_2)
```

# length(unique(sample_ids_new$case_id)) 458

## merge mutation file and sample IDs

```{r merge_mut_sample}
maf_barcode_ids_mutations_sampleIDs <- merge(maf_barcode_ids_mutations, sample_ids_new_case_ids, by.x = "Tumor_Sample_UUID", by.y="aliquot_id")
```


```{r merge_dMDT_file_mut}
old_new_id_match <- merge(history_updated2, tcga_case_id, by.x = 'newid', by.y='file_id')
old_new_id_match2 <- merge(history_updated2, tcga_case_id_rest, by.x = 'oldid', by.y='file_id')
old_new_id_matches <- rbind(old_new_id_match, old_new_id_match2)
```


```{r merge2}
maf_barcode_ids_mutations_sampleIDs_part1 <- maf_barcode_ids_mutations_sampleIDs[maf_barcode_ids_mutations_sampleIDs$file_id %in% old_new_id_matches$oldid,]
maf_barcode_ids_mutations_sampleIDs_part2 <-  maf_barcode_ids_mutations_sampleIDs[maf_barcode_ids_mutations_sampleIDs$file_id %in% old_new_id_matches$newid,]
maf_barcode_ids_mutations_sampleIDs_total <- rbind(maf_barcode_ids_mutations_sampleIDs_part1, maf_barcode_ids_mutations_sampleIDs_part2)
```


```{r merge_dMDT_occurance}
dMDT_Melanoma_tcga <- rbind(dMDT_Melanoma_tcga_metastatis, dMDT_Melanoma_tcga_primary)
dMDT_Melanoma_tcga_fileid <- merge(dMDT_Melanoma_tcga, old_new_id_matches, by.x='SampleID', by.y='oldid')
dMDT_Melanoma_tcga_fileid2 <- merge(dMDT_Melanoma_tcga, old_new_id_matches, by.x='SampleID', by.y='newid')
colnames(dMDT_Melanoma_tcga_fileid2) <- colnames(dMDT_Melanoma_tcga_fileid)
dMDT_Melanoma_tcga_fileid_finale <- rbind(dMDT_Melanoma_tcga_fileid2, dMDT_Melanoma_tcga_fileid)

dMDT_Melanoma_tcga_fileids_mutations <- merge(dMDT_Melanoma_tcga_fileid_finale, maf_barcode_ids_mutations_sampleIDs_total, by.x='newid', by.y='file_id')
```


```{r dMDT_occurance}
dMDT_Melanoma_tcga_correlations <- dMDT_Melanoma_tcga_fileids_mutations %>% dplyr::select(SampleID, case_id, ENSG, dMDT, MDT_GTEx, Frame_Shift_Del, Frame_Shift_Ins, In_Frame_Del, In_Frame_Ins, Missense_Mutation, Nonsense_Mutation, Nonstop_Mutation, Splice_Site, Silent, Translation_Start_Site, Intron, Splice_Region, total) %>% dplyr::distinct()

Mel_dMDT_Occurance_mut <- dMDT_Melanoma_tcga_correlations %>% dplyr::group_by(SampleID, case_id) %>% dplyr::count(dMDT) %>% dplyr::select(SampleID,case_id,n) %>% summarize(total_dMDT = sum(n)) %>% arrange(desc(total_dMDT))
head(Mel_dMDT_Occurance_mut)

dMDT_Melanoma_tcga_correlations_CORR <- merge(dMDT_Melanoma_tcga_correlations, Mel_dMDT_Occurance_mut, by="SampleID")
```


```{r barplot_correlations}
dMDT_Melanoma_tcga_correlations_CORR2 <- dMDT_Melanoma_tcga_correlations_CORR %>% dplyr::select(SampleID, case_id.x, Frame_Shift_Del, Frame_Shift_Ins, In_Frame_Del, In_Frame_Ins, Missense_Mutation, Nonsense_Mutation, Nonstop_Mutation, Splice_Site, Silent, Translation_Start_Site, Intron, Splice_Region, total, total_dMDT) %>% dplyr::distinct()

plot(dMDT_Melanoma_tcga_correlations_CORR2$Frame_Shift_Del, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Frame_Shift_Ins, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$In_Frame_Del, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Missense_Mutation, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Nonsense_Mutation, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Splice_Site, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Silent, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Translation_Start_Site, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$Splice_Region, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
plot(dMDT_Melanoma_tcga_correlations_CORR2$total, dMDT_Melanoma_tcga_correlations_CORR2$total_dMDT)
```


## splice protein mutation?
```{r select_Spliceosome}
maf <- GDCprepare(query)
maf_genes <- maf %>% dplyr::select(Hugo_Symbol, Variant_Classification, Tumor_Sample_UUID, case_id, Gene, Consequence) %>% dplyr::distinct()

Spliceosome_file <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/Final_MDTs_Switches/Spliceosome/idmapping_2023_07_02.tsv', header = TRUE, sep = '\t')
head(Spliceosome_file)

Spliceosome_file$To <- gsub("\\..*", "", Spliceosome_file$To)
ensembl_ids_spliceosome_complex <- unique(Spliceosome_file$To)
maf_spliceosome_genes <- maf_genes[maf_genes$Gene %in% ensembl_ids_spliceosome_complex, ] # ensembl_ids_spliceosome_complex comes from SplicingMachinery Rmd file 
```


```{r merge_samples_spliceosome}
maf_spliceosome_genes_missense_occur <- maf_spliceosome_genes %>% dplyr::filter(Variant_Classification %in% c('Missense_Mutation', "Splice_Site", "3'UTR", "In_Frame_Del", "Frame_Shift_Del", "Splice_Region","5'UTR", "splice_acceptor_variant","Translation_Start_Site", "Nonsense_Mutation","Intron", "Silent")) %>% group_by(Tumor_Sample_UUID)  %>% dplyr::count(Gene) %>% dplyr::select(Tumor_Sample_UUID, n) %>% dplyr::reframe(total_mut = sum(n))

spliceosome_maf_samples <- merge(dMDT_Melanoma_tcga_fileids_mutations, maf_spliceosome_genes_missense_occur, by='Tumor_Sample_UUID')

spliceosome_maf_samples2 <- merge(dMDT_Melanoma_tcga_fileids_mutations, maf_spliceosome_genes, by='Tumor_Sample_UUID')
head(spliceosome_maf_samples2)
```


```{r spliceosomeMutSpliced}
Samples_w_Splicesome_Switch <- unique(dMDT_Melanoma_tcga_correlations[dMDT_Melanoma_tcga_correlations$ENSG %in% ensembl_ids_spliceosome_complex, 'SampleID']) # 86 samples have switch in their spliceosome genes

Samples_w_Spliceosome_Mut <- unique(dMDT_Melanoma_tcga_fileids_mutations[dMDT_Melanoma_tcga_fileids_mutations$Tumor_Sample_UUID %in% maf_spliceosome_genes$Tumor_Sample_UUID, 'SampleID']) # 314 samples have mutation in their spliceosome genes

Samples_w_Spliceosome_Switch_Mut <- intersect(Samples_w_Splicesome_Switch, Samples_w_Spliceosome_Mut)
# 62 unique samples have spliceosome switch and mutation in their spliceosome genes. 

Samples_only_Mut <- setdiff(Samples_w_Spliceosome_Mut, Samples_w_Splicesome_Switch) # 252 samples 
Samples_only_Switch <- setdiff(Samples_w_Splicesome_Switch, Samples_w_Spliceosome_Mut)
```

```{r boxplots}
dMDT_count_Spliceosome_Mut_Switch <- Mel_dMDT_Occurance_mut[Mel_dMDT_Occurance_mut$SampleID %in% Samples_w_Spliceosome_Switch_Mut, ]

dMDT_count_Spliceosome_Mut_Switch$Cohort <- 'Mut_Switch'

dMDT_count_Spliceosome_Only_Mut <- Mel_dMDT_Occurance_mut[Mel_dMDT_Occurance_mut$SampleID %in% Samples_only_Mut, ]
dMDT_count_Spliceosome_Only_Mut$Cohort <- 'Only_Mut'

dMDT_count_Spliceosome_Only_Switch <- Mel_dMDT_Occurance_mut[Mel_dMDT_Occurance_mut$SampleID %in% Samples_only_Switch, ]
dMDT_count_Spliceosome_Only_Switch$Cohort <- 'Only_Switch'

all_samples_switch_mut <- c(dMDT_count_Spliceosome_Only_Switch, dMDT_count_Spliceosome_Only_Mut, dMDT_count_Spliceosome_Mut_Switch)

dMDT_count_rest_of_the_samples <- Mel_dMDT_Occurance_mut[!Mel_dMDT_Occurance_mut$SampleID %in% all_samples_switch_mut, ]
dMDT_count_rest_of_the_samples$Cohort <- 'Others'

dMDT_count_Cohort <- rbind(dMDT_count_Spliceosome_Mut_Switch, dMDT_count_Spliceosome_Only_Mut, dMDT_count_Spliceosome_Only_Switch, dMDT_count_rest_of_the_samples)

ggplot(dMDT_count_Cohort, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot() + theme_classic()
```


wilcox.test(dMDT_count_Cohort[dMDT_count_Cohort$Cohort == 'Only_Switch', 'total_dMDT']$total_dMDT, dMDT_count_Cohort[dMDT_count_Cohort$Cohort == 'Only_Mut', 'total_dMDT']$total_dMDT)

wilcox.test(dMDT_count_Cohort[dMDT_count_Cohort$Cohort == 'Only_Switch', 'total_dMDT']$total_dMDT, dMDT_count_Cohort[dMDT_count_Cohort$Cohort == 'NonMutated', 'total_dMDT']$total_dMDT)


Two samples with PRPF8 missense mutations and have switched isoforms!


311 cases with dMDTs and having mutations
we do not see switch in the mostly mutated genes in melanoma
dMDT_mutated_all[dMDT_mutated_all$Hugo_Symbol %in%  c('TTN', 'MUC16', 'DNAH5', 'PCLO', 'LPR1B', 'ANK3', 'ADGRV1', 'DNAH7', 'CSMD1', 'BRAF'), ]

## Ovarian Samples
```{r maf_df_ovary}
maf_barcode_ids_Ovary <- maf_Ovary %>% dplyr::select(Tumor_Sample_Barcode, Tumor_Sample_UUID) %>% dplyr::distinct()
```

```{r merge_files}
maf_Ovary <- read.maf(maf = maf_Ovary)
maf_df_ovary <- mafSummary(maf_Ovary)$variant.classification.summary
maf_barcode_ids_mutations_ovary <- merge(maf_barcode_ids_Ovary, maf_df_ovary, by= 'Tumor_Sample_Barcode')
```


```{r read_switches}
dMDT_Ovary_tcga_primary <- readRDS('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Ovary/Switches/TCGA_GTEx_ov_Switch.RDS')

```

```{r TCGA_id_convert}
library(TCGAutils)

sample_ids_ovary <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/Ovary/TCGA_SampleInfo/biospecimen.cases_selection.2022-11-17/aliquot.tsv', sep = '\t', header=TRUE)


# convert old TCGA ids to new ids
history_updated2 <- data.frame(oldid = character(), newid = character())
for (eachfile in c(unique(dMDT_Ovary_tcga_primary$SampleID))) {
  updated_list <- UUIDhistory(eachfile)
  oldfile <- updated_list[1,1]
  newfile <- updated_list[2,1]
  history_updated <- data.frame(oldid = oldfile, newid = newfile)
  history_updated2 <- rbind(history_updated2, history_updated)
}

#tcga_case_id_ovary <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$newid,to_type = "case_id")
tcga_case_id_rest_ovary <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$oldid,to_type = "case_id") ## Ovarian ids are still in use in TCGA 
#head(tcga_case_id_ovary)
head(tcga_case_id_rest_ovary)
```


```{r merge_sample_id}
sample_ids_new <- rbind(sample_ids_ovary[sample_ids_ovary$case_id %in% tcga_case_id_rest_ovary$cases.case_id,], sample_ids_ovary[sample_ids_ovary$case_id %in% tcga_case_id_rest_ovary$file_id,])

sample_ids_new_case_ids_ovary <- merge(sample_ids_new, tcga_case_id_rest_ovary, by.x= "case_id", by.y="cases.case_id")
```


```{r merge_mut_sample}
maf_barcode_ids_mutations_sampleIDs_ovary <- merge(maf_barcode_ids_mutations_ovary, sample_ids_new_case_ids_ovary, by.x = "Tumor_Sample_UUID", by.y="aliquot_id")
```


```{r merge_dMDT_file_mut}
#old_new_id_match <- merge(history_updated2, tcga_case_id, by.x = 'newid', by.y='file_id')
old_new_id_matches <- merge(history_updated2, tcga_case_id_rest_ovary, by.x = 'oldid', by.y='file_id')
```


```{r merge_OVARY}
maf_barcode_ids_mutations_sampleIDs_total_ovary <-  maf_barcode_ids_mutations_sampleIDs_ovary[maf_barcode_ids_mutations_sampleIDs_ovary$file_id %in% old_new_id_matches$oldid,]
```

```{r select_Spliceosome}
maf_Ovary <- GDCprepare(query_Ovary)
maf_genes_Ovary <- maf_Ovary %>% dplyr::select(Hugo_Symbol, Variant_Classification, Tumor_Sample_UUID, case_id, Gene, Consequence) %>% dplyr::distinct()

maf_spliceosome_genes_Ovary <- maf_genes_Ovary[maf_genes_Ovary$Gene %in% ensembl_ids_spliceosome_complex, ] # ensembl_ids_spliceosome_complex comes from SplicingMachinery Rmd file 
```

```{r merge_dMDT_occurance}
dMDT_Ovary_tcga_fileid <- merge(dMDT_Ovary_tcga_primary, old_new_id_matches, by.x='SampleID', by.y='oldid')
dMDT_Ovary_tcga_fileids_mutations <- merge(dMDT_Ovary_tcga_fileid, maf_barcode_ids_mutations_sampleIDs_total_ovary, by.x='SampleID', by.y='file_id')
```


```{r dMDT_occuranceOvary}
dMDT_Ovary_tcga_correlations <- dMDT_Ovary_tcga_fileids_mutations %>% dplyr::select(SampleID, case_id, ENSG, dMDT, MDT_GTEx, Frame_Shift_Del, Frame_Shift_Ins, In_Frame_Del, In_Frame_Ins, Missense_Mutation, Nonsense_Mutation, Nonstop_Mutation, Splice_Site, Silent, Translation_Start_Site, Intron, Splice_Region, total) %>% dplyr::distinct()

Ov_dMDT_Occurance_mut <- dMDT_Ovary_tcga_correlations %>% dplyr::group_by(SampleID, case_id) %>% dplyr::count(dMDT) %>% dplyr::select(SampleID,case_id,n) %>% summarize(total_dMDT = sum(n)) %>% arrange(desc(total_dMDT))
head(Ov_dMDT_Occurance_mut)

#dMDT_Ov_tcga_correlations_CORR <- merge(dMDT_Ovary_tcga_correlations, Ov_dMDT_Occurance_mut, by="SampleID")
```

```{r spliceosomeMutSpliced}
Samples_w_Splicesome_Switch_Ov <- unique(dMDT_Ovary_tcga_fileids_mutations[dMDT_Ovary_tcga_fileids_mutations$ENSG %in% ensembl_ids_spliceosome_complex, 'SampleID']) # 86 samples have switch in their spliceosome genes

Samples_w_Spliceosome_Mut_Ov <- unique(dMDT_Ovary_tcga_fileids_mutations[dMDT_Ovary_tcga_fileids_mutations$Tumor_Sample_UUID %in% dMDT_Ovary_tcga_fileids_mutations$Tumor_Sample_UUID, 'SampleID']) # 314 samples have mutation in their spliceosome genes

Samples_w_Spliceosome_Switch_Mut_Ov <- intersect(Samples_w_Splicesome_Switch_Ov, Samples_w_Spliceosome_Mut_Ov)
# 62 unique samples have spliceosome switch and mutation in their spliceosome genes. 

Samples_only_Mut <- setdiff(Samples_w_Spliceosome_Mut_Ov, Samples_w_Splicesome_Switch_Ov) # 252 samples 
Samples_only_Switch <- setdiff(Samples_w_Splicesome_Switch_Ov, Samples_w_Spliceosome_Mut_Ov)
```

```{r boxplots}
dMDT_count_Spliceosome_Mut_Switch_Ov <- Ov_dMDT_Occurance_mut[Ov_dMDT_Occurance_mut$SampleID %in% Samples_w_Spliceosome_Switch_Mut_Ov, ]

dMDT_count_Spliceosome_Mut_Switch_Ov$Cohort <- 'Mut_Switch'

dMDT_count_Spliceosome_Only_Mut_Ov <- Ov_dMDT_Occurance_mut[Ov_dMDT_Occurance_mut$SampleID %in% Samples_only_Mut, ]
dMDT_count_Spliceosome_Only_Mut_Ov$Cohort <- 'Only_Mut'

dMDT_count_Spliceosome_Only_Switch_Ov <- Ov_dMDT_Occurance_mut[Ov_dMDT_Occurance_mut$SampleID %in% Samples_only_Switch, ]
dMDT_count_Spliceosome_Only_Switch_Ov$Cohort <- 'Only_Switch'

all_samples_switch_mut_Ov <- c(dMDT_count_Spliceosome_Only_Switch_Ov, dMDT_count_Spliceosome_Only_Mut_Ov, dMDT_count_Spliceosome_Mut_Switch_Ov)

dMDT_count_rest_of_the_samples_Ov <- Ov_dMDT_Occurance_mut[!Ov_dMDT_Occurance_mut$SampleID %in% all_samples_switch_mut, ]
dMDT_count_rest_of_the_samples_Ov$Cohort <- 'Others'

dMDT_count_Cohort_Ov <- rbind(dMDT_count_Spliceosome_Mut_Switch_Ov, dMDT_count_Spliceosome_Only_Mut_Ov, dMDT_count_Spliceosome_Only_Switch_Ov, dMDT_count_rest_of_the_samples_Ov)

ggplot(dMDT_count_Cohort_Ov, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot() + theme_classic()
```


## AML Samples
```{r maf_df_AML}
maf_barcode_ids_AML <- maf_AML %>% dplyr::select(Tumor_Sample_Barcode, Tumor_Sample_UUID) %>% dplyr::distinct()
```

```{r merge_files}
maf_AML <- read.maf(maf = maf_AML)
maf_df_AML <- mafSummary(maf_AML)$variant.classification.summary
maf_barcode_ids_mutations_AML <- merge(maf_barcode_ids_AML, maf_df_AML, by= 'Tumor_Sample_Barcode')
```

```{r read_switches}
dMDT_AML_tcga_primary <- readRDS('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/AML/Switches/Switches_TCGA_GTEx.RDS')

```

```{r TCGA_id_convert}
library(TCGAutils)

sample_ids_AML <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/AML/biospecimen.cases_selection.2023-04-18/aliquot.tsv', sep = '\t', header=TRUE)


# convert old TCGA ids to new ids
history_updated2 <- data.frame(oldid = character(), newid = character())
for (eachfile in c(unique(dMDT_AML_tcga_primary$SampleID))) {
  updated_list <- UUIDhistory(eachfile)
  oldfile <- updated_list[1,1]
  newfile <- updated_list[2,1]
  history_updated <- data.frame(oldid = oldfile, newid = newfile)
  history_updated2 <- rbind(history_updated2, history_updated)
}

#tcga_case_id_AML <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$newid,to_type = "case_id")
tcga_case_id_rest_AML <- UUIDtoUUID(history_updated2[2:nrow(history_updated2),]$oldid,to_type = "case_id") ## AML ids are still in use in TCGA 
#head(tcga_case_id_AML)
head(tcga_case_id_rest_AML)
```


```{r merge_sample_id}
sample_ids_new <- rbind(sample_ids_AML[sample_ids_AML$case_id %in% tcga_case_id_rest_AML$cases.case_id,], sample_ids_AML[sample_ids_AML$case_id %in% tcga_case_id_rest_AML$file_id,])

sample_ids_new_case_ids_AML <- merge(sample_ids_new, tcga_case_id_rest_AML, by.x= "case_id", by.y="cases.case_id")
```


```{r merge_mut_sample}
maf_barcode_ids_mutations_sampleIDs_AML <- merge(maf_barcode_ids_mutations_AML, sample_ids_new_case_ids_AML, by.x = "Tumor_Sample_UUID", by.y="aliquot_id")
```


```{r merge_dMDT_file_mut}
#old_new_id_match <- merge(history_updated2, tcga_case_id, by.x = 'newid', by.y='file_id')
old_new_id_matches <- merge(history_updated2, tcga_case_id_rest_AML, by.x = 'oldid', by.y='file_id')
```


```{r merge_AML}
maf_barcode_ids_mutations_sampleIDs_total_AML <-  maf_barcode_ids_mutations_sampleIDs_AML[maf_barcode_ids_mutations_sampleIDs_AML$file_id %in% old_new_id_matches$oldid,]
```

```{r select_Spliceosome}
maf_AML <- GDCprepare(query_AML)
maf_genes_AML <- maf_AML %>% dplyr::select(Hugo_Symbol, Variant_Classification, Tumor_Sample_UUID, case_id, Gene, Consequence) %>% dplyr::distinct()

maf_spliceosome_genes_AML <- maf_genes_AML[maf_genes_AML$Gene %in% ensembl_ids_spliceosome_complex, ] # ensembl_ids_spliceosome_complex comes from SplicingMachinery Rmd file 
```

```{r merge_dMDT_occurance}
dMDT_AML_tcga_fileid <- merge(dMDT_AML_tcga_primary, old_new_id_matches, by.x='SampleID', by.y='oldid')
dMDT_AML_tcga_fileids_mutations <- merge(dMDT_AML_tcga_fileid, maf_barcode_ids_mutations_sampleIDs_total_AML, by.x='SampleID', by.y='file_id')
```


```{r dMDT_occuranceAML}
dMDT_AML_tcga_correlations <- dMDT_AML_tcga_fileids_mutations %>% dplyr::select(SampleID, case_id, ENSG, dMDT, MDT_GTEx, Frame_Shift_Del, Frame_Shift_Ins, In_Frame_Del, In_Frame_Ins, Missense_Mutation, Nonsense_Mutation, Nonstop_Mutation, Splice_Site, Silent, Translation_Start_Site, Intron, Splice_Region, total) %>% dplyr::distinct()

AML_dMDT_Occurance_mut <- dMDT_AML_tcga_correlations %>% dplyr::group_by(SampleID, case_id) %>% dplyr::count(dMDT) %>% dplyr::select(SampleID,case_id,n) %>% summarize(total_dMDT = sum(n)) %>% arrange(desc(total_dMDT))
head(AML_dMDT_Occurance_mut)
#dMDT_AML_tcga_correlations_CORR <- merge(dMDT_AML_tcga_correlations, AML_dMDT_Occurance_mut, by="SampleID")
```

```{r spliceosomeMutSpliced}
Samples_w_Splicesome_Switch_AML <- unique(dMDT_AML_tcga_fileids_mutations[dMDT_AML_tcga_fileids_mutations$ENSG %in% ensembl_ids_spliceosome_complex, 'SampleID']) # 86 samples have switch in their spliceosome genes

Samples_w_Spliceosome_Mut_AML <- unique(dMDT_AML_tcga_fileids_mutations[dMDT_AML_tcga_fileids_mutations$Tumor_Sample_UUID %in% dMDT_AML_tcga_fileids_mutations$Tumor_Sample_UUID, 'SampleID']) # 314 samples have mutation in their spliceosome genes

Samples_w_Spliceosome_Switch_Mut_AML <- intersect(Samples_w_Splicesome_Switch_AML, Samples_w_Spliceosome_Mut_AML)
# 62 unique samples have spliceosome switch and mutation in their spliceosome genes. 

Samples_only_Mut <- setdiff(Samples_w_Spliceosome_Mut_AML, Samples_w_Splicesome_Switch_AML) # 252 samples 
Samples_only_Switch <- setdiff(Samples_w_Splicesome_Switch_AML, Samples_w_Spliceosome_Mut_AML)
```

```{r boxplots}
dMDT_count_Spliceosome_Mut_Switch_AML <- AML_dMDT_Occurance_mut[AML_dMDT_Occurance_mut$SampleID %in% Samples_w_Spliceosome_Switch_Mut_AML, ]

dMDT_count_Spliceosome_Mut_Switch_AML$Cohort <- 'Mut_Switch'

dMDT_count_Spliceosome_Only_Mut_AML <- AML_dMDT_Occurance_mut[AML_dMDT_Occurance_mut$SampleID %in% Samples_only_Mut, ]
dMDT_count_Spliceosome_Only_Mut_AML$Cohort <- 'Only_Mut'

dMDT_count_Spliceosome_Only_Switch_AML <- AML_dMDT_Occurance_mut[AML_dMDT_Occurance_mut$SampleID %in% Samples_only_Switch, ]
dMDT_count_Spliceosome_Only_Switch_AML$Cohort <- 'Only_Switch'

all_samples_switch_mut_AML <- c(dMDT_count_Spliceosome_Only_Switch_AML, dMDT_count_Spliceosome_Only_Mut_AML, dMDT_count_Spliceosome_Mut_Switch_AML)

dMDT_count_rest_of_the_samples_AML <- AML_dMDT_Occurance_mut[!AML_dMDT_Occurance_mut$SampleID %in% all_samples_switch_mut, ]
dMDT_count_rest_of_the_samples_AML$Cohort <- 'Others'

dMDT_count_Cohort_AML <- rbind(dMDT_count_Spliceosome_Mut_Switch_AML, dMDT_count_Spliceosome_Only_Mut_AML, dMDT_count_Spliceosome_Only_Switch_AML, dMDT_count_rest_of_the_samples_AML)

ggplot(dMDT_count_Cohort_AML, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot() + theme_classic()
```

```{r testSignificance}
compare_means(total_dMDT ~ Cohort,  data = dMDT_count_Cohort)
compare_means(total_dMDT ~ Cohort,  data = dMDT_count_Cohort_Ov)
compare_means(total_dMDT ~ Cohort,  data = dMDT_count_Cohort_AML)
```

```{r addSignificance}
comparison_Mel <- list(c("Mut_Switch", "Only_Mut"), c("Mut_Switch", "Only_Switch"), c("Mut_Switch", "Others"), c("Only_Mut", "Only_Switch"), c("Only_Mut", "Others"), c("Only_Switch", "Others"))

comparison_Ov <- list(c("Mut_Switch", "Only_Mut"), c("Mut_Switch", "Others"), c("Only_Mut", "Others"))

comparison_AML <- list(c("Mut_Switch", "Only_Mut"), c("Mut_Switch", "Others"), c("Only_Mut", "Others"))
```

```{r savegraphs}
library(ggpubr)
Spliceosome_Mut_Switch_Mel <- ggplot(dMDT_count_Cohort, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot(notch = FALSE) + theme_classic() + scale_fill_brewer(palette="Blues") + stat_compare_means(comparisons = comparison_Mel)

Spliceosome_Mut_Switch_AML <- ggplot(dMDT_count_Cohort_AML, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot(notch = FALSE) + theme_classic() + scale_fill_brewer(palette="OrRd") + stat_compare_means(comparisons = comparison_AML)

Spliceosome_Mut_Switch_Ov <- ggplot(dMDT_count_Cohort_Ov, aes(x=Cohort, y=total_dMDT, fill=Cohort)) + geom_boxplot(notch = FALSE) + theme_classic() + scale_fill_brewer(palette="RdPu") + stat_compare_means(comparisons = comparison_Ov)


ggsave("/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Figures/Spliceosome_Mut_Switch_Mel.png", plot = Spliceosome_Mut_Switch_Mel, bg = "transparent", width = 10, height = 8)
ggsave("/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Figures/Spliceosome_Mut_Switch_AML.png", plot = Spliceosome_Mut_Switch_AML, bg = "transparent", width = 10, height = 8)
ggsave("/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Figures/Spliceosome_Mut_Switch_Ov.png", plot = Spliceosome_Mut_Switch_Ov, bg = "transparent", width = 10, height = 8)
```




