---
title: "Manuscript File 1"
date: "2025-02-27"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    social: menu
    source: embed
---

This file includes manuscript figures! Figure 2, Table 1, Supplementary Table 1, 2, and 3. Also transmembrane protein statistics. 

```{r libraries, message=FALSE}
library(patchwork)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(knitr)
library(rmarkdown)
library(harmony)
library(Matrix)
library(cowplot)
library(RColorBrewer)
library(viridis)
library(ggpubr)
set.seed(12345)
```


# Check Redundant Transcript Number
```{r checkRedundantTranscriptNumber}
tuPro_primary_tpm <- readRDS('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Ovary/TPMValues/RedundantTPMs/tuPro_tpm_NR_primary.RDS')

tuPro_primary_tpm_redundant <- read.csv('/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Ovary/combined_columns.tsv', sep = '\t', header = TRUE, check.names = FALSE)

dim(tuPro_primary_tpm_redundant)[1] - dim(tuPro_primary_tpm)[1] # 507 first column in the redundant one is tpm
```
length(unique(tuPro_primary_tpm[tuPro_primary_tpm$ENSG %in% biomart_gene_types[biomart_gene_types$Gene.type == 'protein_coding', 'Gene.stable.ID'],'ENSG']))

length(unique(tuPro_primary_tpm[tuPro_primary_tpm$ENSG %in% biomart_gene_types[biomart_gene_types$Gene.type == 'protein_coding', 'Gene.stable.ID'],'ENSG']))

# Figure 2: MDT, MDT Switches (cMDTs) and Overlaps among datasets
## MDT Analysis Across All Datasets
```{r MDT_Numbers, warning=FALSE, message=FALSE}
# ovary
MDTs_Ovary_GTEx <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Ovary_OvarianCancer/GTEx_MDTs_ov_for_comparison_2.RDS')
MDTs_Ovary_TCGA <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Ovary_OvarianCancer/tcga_MDTs_ov.RDS')
MDTs_Ovary_TuPro_primary <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Ovary_OvarianCancer/tuPro_MDTs_ov_primary.RDS')
MDTs_Ovary_TuPro_metastatic <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Ovary_OvarianCancer/tuPro_MDTs_ov_metastatic.RDS')

# melanoma
MDTs_Melanoma_GTEx <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Skin_Melanoma/GTEx_MDTs_mel_2.RDS')
MDTs_Melanoma_TCGA_primary <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Skin_Melanoma/tcga_primary_MDTs_mel.RDS')
MDTs_Melanoma_TCGA_metas <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Skin_Melanoma/tcga_metastatic_MDTs_mel.RDS')
MDTs_Melanoma_TuPro <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Skin_Melanoma/tuPro_MDTs_mel.RDS')

# aml
MDTs_AML_GTEx <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Blood_AML/GTEx_MDTs_AML_for_comparison_2.RDS')
MDTs_AML_TCGA <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Blood_AML/tcga_MDTs_AML.RDS')
MDTs_AML_TuPro <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/MDT/Blood_AML/tuPro_MDTs_AML.RDS')
```

## Figure 2A Data prep
```{r VisualizeMDTs, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
# Ovary
MDTs_GTEx_Ovary_n_per_sample <- MDTs_Ovary_GTEx %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Ovary_n_per_sample <- MDTs_Ovary_TCGA %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Ovary_pri_n_per_sample <- MDTs_Ovary_TuPro_primary %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Ovary_met_n_per_sample <- MDTs_Ovary_TuPro_metastatic %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

# Melanoma
MDTs_GTEx_Mel_n_per_sample <- MDTs_Melanoma_GTEx %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Mel_pri_n_per_sample <- MDTs_Melanoma_TCGA_primary %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Mel_met_n_per_sample <- MDTs_Melanoma_TCGA_metas %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Mel_n_per_sample <- MDTs_Melanoma_TuPro %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

# AML
MDTs_GTEx_AML_n_per_sample <- MDTs_AML_GTEx %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_tcga_AML_n_per_sample <- MDTs_AML_TCGA %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_AML_n_per_sample <- MDTs_AML_TuPro %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 


MDT_counts_per_samples <- rbind(cbind(type = rep('GTEx_Mel', dim(MDTs_GTEx_Mel_n_per_sample)[1]),      
                                      MDTs_GTEx_Mel_n_per_sample[,1:2]), 
                                cbind(type = rep('TCGA_pri_Mel', dim(MDTs_TCGA_Mel_pri_n_per_sample)[1]),     MDTs_TCGA_Mel_pri_n_per_sample[,1:2]),
                                   cbind(type = rep('TCGA_Mel_Metastatic', dim(MDTs_TCGA_Mel_met_n_per_sample)[1]), MDTs_TCGA_Mel_met_n_per_sample[,1:2]),
                        cbind(type = rep('TP_Mel', dim(MDTs_TuPro_Mel_n_per_sample)[1]), MDTs_TuPro_Mel_n_per_sample[,1:2]),
                        cbind(type = rep('GTEx_Ov', dim(MDTs_GTEx_Ovary_n_per_sample)[1]), MDTs_GTEx_Ovary_n_per_sample[,1:2]),
                        cbind(type = rep('TCGA_Ov', dim(MDTs_TCGA_Ovary_n_per_sample)[1]), MDTs_TCGA_Ovary_n_per_sample[,1:2]), 
                        cbind(type = rep('TP_Ov', dim(MDTs_TuPro_Ovary_pri_n_per_sample)[1]), MDTs_TuPro_Ovary_pri_n_per_sample[,1:2]), 
                        cbind(type = rep('TP_Ov_Metastatic', dim(MDTs_TuPro_Ovary_met_n_per_sample)[1]), MDTs_TuPro_Ovary_met_n_per_sample[,1:2]), cbind(type = rep('GTEx_AML', dim(MDTs_GTEx_AML_n_per_sample)[1]), MDTs_GTEx_AML_n_per_sample[,1:2]), 
                        cbind(type = rep('TCGA_AML', dim(MDTs_tcga_AML_n_per_sample)[1]), MDTs_tcga_AML_n_per_sample[,1:2]),cbind(type = rep('TP_AML', dim(MDTs_TuPro_AML_n_per_sample)[1]), MDTs_TuPro_AML_n_per_sample[,1:2]))

library(ggplot2)
library(RColorBrewer)

MDT_counts_per_samples$type <- factor(MDT_counts_per_samples$type, levels = unique(MDT_counts_per_samples$type))

# If you have 3 main types, you can choose a palette of 3 colors.
my_palette <- c("#d85bd4",
                "#db95d9",
                "#dcb7dc",
                "#2d858d",
                "#34cacd",
                "#73e5ea",
                "#df8250",
                "#e4a481")


# Interpolate to create a palette with 11 colors
expanded_palette <- colorRampPalette(my_palette)(11)
```

## Figure 2A
```{r NumberOfSamples}
NumberOfSamples <- data.frame(Cohort = c('GTEx Skin', 'TCGA Primary Melanoma', 'TCGA Metastatic Melanoma', 'TuPro Melanoma', 
                                         'GTEx Ovary', 'TCGA Ovary', 'TuPro Primary Ovary', 'TuPro Metastatic Ovary', 
                                         'GTEx Blood', 'TCGA AML', 'TuPro AML'), 
                              nSample = c(length(unique(MDTs_Melanoma_GTEx$SampleID)), length(unique(MDTs_Melanoma_TCGA_primary$SampleID)), 
                                          length(unique(MDTs_Melanoma_TCGA_metas$SampleID)), 
                                          length(unique(MDTs_Melanoma_TuPro$SampleID)), 
                                          length(unique(MDTs_Ovary_GTEx$SampleID)), 
                                          length(unique(MDTs_Ovary_TCGA$SampleID)), 
                                          length(unique(MDTs_Ovary_TuPro_primary$SampleID)), 
                                          length(unique(MDTs_Ovary_TuPro_metastatic$SampleID)), 
                                          length(unique(MDTs_AML_GTEx$SampleID)), 
                                          length(unique(MDTs_AML_TCGA$SampleID)), 
                                          length(unique(MDTs_AML_TuPro$SampleID))))

head(NumberOfSamples)
NumberOfSamples$Cohort <- factor(NumberOfSamples$Cohort, levels = unique(NumberOfSamples$Cohort))

my_palette <- c("#d85bd4",
                "#db95d9",
                "#dcb7dc",
                "#2d858d",
                "#34cacd",
                "#73e5ea",
                "#df8250",
                "#e4a481")
expanded_palette <- colorRampPalette(my_palette)(11)
a <- ggplot(NumberOfSamples, aes(x=Cohort, y = nSample, fill = Cohort)) +
    geom_bar(stat = 'identity') +
    scale_fill_manual(values = expanded_palette)  +
    theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    ylim(0, 500)

a + theme(axis.text.x = element_blank(), legend.text = element_blank()) 
```

### Total number of genes with MDTs across samples
```{r MDTs}
means_df <- MDT_counts_per_samples %>%
  group_by(type) %>%
  summarise(mean_value = mean(n))

mean(means_df$mean_value)
#[1] 6626.748 on average 

# Find the maximum y value to place the text above everything else
max_y <- max(MDT_counts_per_samples$n, na.rm = TRUE)

p <- ggplot(MDT_counts_per_samples, aes(x = type, y = n, color = type)) +
    geom_boxplot() +
    geom_text(data = means_df, aes(x = type, y = max_y+10, label = round(mean_value, 1)), 
            position = position_nudge(y = 500), vjust = 0, hjust = 1, size = 3, color = "darkgray", angle=90) +
    scale_color_manual(values = expanded_palette) + 
    xlab("SampleID") + 
    ylab("Number of MDTs") +
    ggtitle("Distribution of Number of MDTs per SampleID") + theme_classic() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.margin = unit(c(1,1,1,1), "lines"))

p + theme(axis.text.x = element_blank(), legend.text = element_blank()) 

```

### Total number of Genes with a MDT -  it is identified for almost every gene. 
```{r checkNumberOfMDTs}
# number of MDTs
length(unique(c(MDTs_Ovary_GTEx$ENST1,
MDTs_Ovary_TCGA$ENST1,
MDTs_Ovary_TuPro_primary$ENST1,
MDTs_Ovary_TuPro_metastatic$ENST1,
MDTs_Melanoma_GTEx$ENST1,
MDTs_Melanoma_TCGA_primary$ENST1,
MDTs_Melanoma_TCGA_metas$ENST1,
MDTs_Melanoma_TuPro$ENST1,
MDTs_AML_GTEx$ENST1,
MDTs_AML_TCGA$ENST1,
MDTs_AML_TuPro$ENST1)))

# number of genes having MDTs
length(unique(c(MDTs_Ovary_GTEx$ENSG,
MDTs_Ovary_TCGA$ENSG,
MDTs_Ovary_TuPro_primary$ENSG,
MDTs_Ovary_TuPro_metastatic$ENSG,
MDTs_Melanoma_GTEx$ENSG,
MDTs_Melanoma_TCGA_primary$ENSG,
MDTs_Melanoma_TCGA_metas$ENSG,
MDTs_Melanoma_TuPro$ENSG,
MDTs_AML_GTEx$ENSG,
MDTs_AML_TCGA$ENSG,
MDTs_AML_TuPro$ENSG)))
```


### Average number of genes with MDT 
```{r MeanOfMDTs}
MDT_counts_per_samples %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n))
# total mean across all cohorts
MDT_counts_per_samples %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n)) %>% dplyr::select(mean) %>% dplyr::summarise(mean2 = mean(mean))

# avg of cancer samples
MDT_counts_per_samples %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n)) %>% dplyr::filter(type %in% c('TCGA_pri_Mel', 'TCGA_Mel_Metastatic', 'TP_Mel', 'TCGA_Ov', 'TP_Ov', 'TP_Ov_Metastatic', 'TCGA_AML')) %>% dplyr::select(mean) %>% dplyr::summarise(mean2 = mean(mean))

# avg of gtex samples
MDT_counts_per_samples %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n)) %>% dplyr::filter(type %in% c('GTEx_Mel', 'GTEx_Ov', 'GTEx_AML')) %>% dplyr::select(mean) %>% dplyr::summarise(mean2 = mean(mean))
```


```{r distributionOfRates, warning=FALSE, message=FALSE}
# check distribution of rates - how many times MDT1 is higher than the MDT2 in normal and cancer samples
MDT_rates_per_samples <- rbind(cbind(type = rep('GTEx_Mel', dim(MDTs_Melanoma_GTEx)[1]),      
                                      MDTs_Melanoma_GTEx[,c(1,7)]), 
                                cbind(type = rep('TCGA_pri_Mel', dim(MDTs_Melanoma_TCGA_primary)[1]),     MDTs_Melanoma_TCGA_primary[,c(1,7)]),
                                   cbind(type = rep('TCGA_Mel_Metastatic', dim(MDTs_Melanoma_TCGA_metas)[1]), MDTs_Melanoma_TCGA_metas[,c(1,7)]),
                        cbind(type = rep('TP_Mel', dim(MDTs_Melanoma_TuPro)[1]), MDTs_Melanoma_TuPro[,c(1,7)]),
                        cbind(type = rep('GTEx_Ov', dim(MDTs_Ovary_GTEx)[1]), MDTs_Ovary_GTEx[,c(1,7)]),
                        cbind(type = rep('TCGA_Ov', dim(MDTs_Ovary_TCGA)[1]), MDTs_Ovary_TCGA[,c(1,7)]), 
                        cbind(type = rep('TP_Ov', dim(MDTs_Ovary_TuPro_primary)[1]), MDTs_Ovary_TuPro_primary[,c(1,7)]), 
                        cbind(type = rep('TP_Ov_Metastatic', dim(MDTs_Ovary_TuPro_metastatic)[1]), MDTs_Ovary_TuPro_metastatic[,c(1,7)]), cbind(type = rep('GTEx_AML', dim(MDTs_AML_GTEx)[1]), MDTs_AML_GTEx[,c(1,7)]), 
                        cbind(type = rep('TCGA_AML', dim(MDTs_AML_TCGA)[1]), MDTs_AML_TCGA[,c(1,7)]),cbind(type = rep('TP_AML', dim(MDTs_AML_TuPro)[1]), MDTs_AML_TuPro[,c(1,7)]))
```

median(MDT_rates_per_samples$rate) # 5.44

```{r ProteinCoding, warning=FALSE, message=FALSE}
biomart_gene_types <- read.csv('/Volumes/PersonalDisk/PhD_Macbook/PhD/Projects/TumorProfiler/Analysis/ensg_genetype.txt', sep='\t', header = TRUE)
dim(biomart_gene_types)
#[1] 67128     2
gene_names_pc <- biomart_gene_types[biomart_gene_types$Gene.type == 'protein_coding', 1]
length(unique(biomart_gene_types[biomart_gene_types$Gene.type == 'protein_coding', 1]))
```

## Figure 2B - distribution of MDTs across samples
```{r isolateProteinCodingMDTs, warning=FALSE, message=FALSE}
# ovary
MDTs_Ovary_GTEx_coding <- MDTs_Ovary_GTEx[MDTs_Ovary_GTEx$ENSG %in% gene_names_pc, ]
MDTs_Ovary_TCGA_coding <- MDTs_Ovary_TCGA[MDTs_Ovary_TCGA$ENSG %in% gene_names_pc, ]
MDTs_Ovary_TuPro_primary_coding <- MDTs_Ovary_TuPro_primary[MDTs_Ovary_TuPro_primary$ENSG %in% gene_names_pc, ]
MDTs_Ovary_TuPro_metastatic_coding <- MDTs_Ovary_TuPro_metastatic[MDTs_Ovary_TuPro_metastatic$ENSG %in% gene_names_pc, ]

# melanoma
MDTs_Melanoma_GTEx_coding <- MDTs_Melanoma_GTEx[MDTs_Melanoma_GTEx$ENSG %in% gene_names_pc, ]
MDTs_Melanoma_TCGA_primary_coding <- MDTs_Melanoma_TCGA_primary[MDTs_Melanoma_TCGA_primary$ENSG  %in% gene_names_pc, ]
MDTs_Melanoma_TCGA_metas_coding <- MDTs_Melanoma_TCGA_metas[MDTs_Melanoma_TCGA_metas$ENSG %in% gene_names_pc, ]
MDTs_Melanoma_TuPro_coding <- MDTs_Melanoma_TuPro[MDTs_Melanoma_TuPro$ENSG %in% gene_names_pc, ]

# aml
MDTs_AML_GTEx_coding <- MDTs_AML_GTEx[MDTs_AML_GTEx$ENSG %in% gene_names_pc, ]
MDTs_AML_TCGA_coding <-MDTs_AML_TCGA[MDTs_AML_TCGA$ENSG %in% gene_names_pc, ]
MDTs_AML_TuPro_coding <- MDTs_AML_TuPro[MDTs_AML_TuPro$ENSG %in% gene_names_pc, ]


# Ovary
MDTs_GTEx_Ovary_n_per_sample_coding <- MDTs_Ovary_GTEx_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Ovary_n_per_sample_coding <- MDTs_Ovary_TCGA_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Ovary_pri_n_per_sample_coding <- MDTs_Ovary_TuPro_primary_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Ovary_met_n_per_sample_coding <- MDTs_Ovary_TuPro_metastatic_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

# Melanoma
MDTs_GTEx_Mel_n_per_sample_coding <- MDTs_Melanoma_GTEx_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Mel_pri_n_per_sample_coding <- MDTs_Melanoma_TCGA_primary_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TCGA_Mel_met_n_per_sample_coding <- MDTs_Melanoma_TCGA_metas_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_Mel_n_per_sample_coding <- MDTs_Melanoma_TuPro_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

# AML
MDTs_GTEx_AML_n_per_sample_coding <- MDTs_AML_GTEx_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_tcga_AML_n_per_sample_coding <- MDTs_AML_TCGA_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

MDTs_TuPro_AML_n_per_sample_coding <- MDTs_AML_TuPro_coding %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 


MDT_counts_per_samples_coding <- rbind(cbind(type = rep('GTEx_Mel', dim(MDTs_GTEx_Mel_n_per_sample_coding)[1]),      
                                      MDTs_GTEx_Mel_n_per_sample_coding[,1:2]), 
                                cbind(type = rep('TCGA_pri_Mel', dim(MDTs_TCGA_Mel_pri_n_per_sample_coding)[1]),     MDTs_TCGA_Mel_pri_n_per_sample_coding[,1:2]),
                                   cbind(type = rep('TCGA_Mel_Metastatic', dim(MDTs_TCGA_Mel_met_n_per_sample_coding)[1]), MDTs_TCGA_Mel_met_n_per_sample_coding[,1:2]),
                        cbind(type = rep('TP_Mel', dim(MDTs_TuPro_Mel_n_per_sample_coding)[1]), MDTs_TuPro_Mel_n_per_sample_coding[,1:2]),
                        cbind(type = rep('GTEx_Ov', dim(MDTs_GTEx_Ovary_n_per_sample_coding)[1]), MDTs_GTEx_Ovary_n_per_sample_coding[,1:2]),
                        cbind(type = rep('TCGA_Ov', dim(MDTs_TCGA_Ovary_n_per_sample_coding)[1]), MDTs_TCGA_Ovary_n_per_sample_coding[,1:2]), 
                        cbind(type = rep('TP_Ov', dim(MDTs_TuPro_Ovary_pri_n_per_sample_coding)[1]), MDTs_TuPro_Ovary_pri_n_per_sample_coding[,1:2]), 
                        cbind(type = rep('TP_Ov_Metastatic', dim(MDTs_TuPro_Ovary_met_n_per_sample_coding)[1]), MDTs_TuPro_Ovary_met_n_per_sample_coding[,1:2]), cbind(type = rep('GTEx_AML', dim(MDTs_GTEx_AML_n_per_sample_coding)[1]), MDTs_GTEx_AML_n_per_sample_coding[,1:2]), 
                        cbind(type = rep('TCGA_AML', dim(MDTs_tcga_AML_n_per_sample_coding)[1]), MDTs_tcga_AML_n_per_sample_coding[,1:2]),cbind(type = rep('TP_AML', dim(MDTs_TuPro_AML_n_per_sample_coding)[1]), MDTs_TuPro_AML_n_per_sample_coding[,1:2]))

library(ggplot2)
library(RColorBrewer)

MDT_counts_per_samples_coding$type <- factor(MDT_counts_per_samples_coding$type, levels = unique(MDT_counts_per_samples_coding$type))

# If you have 3 main types, you can choose a palette of 3 colors.
my_palette <- c("#d85bd4",
                "#db95d9",
                "#dcb7dc",
                "#2d858d",
                "#34cacd",
                "#73e5ea",
                "#df8250",
                "#e4a481")


# Interpolate to create a palette with 11 colors
expanded_palette <- colorRampPalette(my_palette)(11)



means_df_coding <- MDT_counts_per_samples_coding %>%
  group_by(type) %>%
  summarise(mean_value = mean(n))

mean(means_df$mean_value)
#[1] 6626.748 on average 

p_coding <- ggplot(MDT_counts_per_samples_coding, aes(x = type, y = n, color = type)) +
    geom_boxplot() +
    geom_text(data = means_df_coding, aes(x = type, y = max_y+10, label = round(mean_value, 1)), 
            position = position_nudge(y = 500), vjust = 0, hjust = 1, size = 3, color = "darkgray", angle=90) +
    scale_color_manual(values = expanded_palette) + 
    xlab("SampleID") + 
    ylab("Number of MDTs") +
    ggtitle("Distribution of Number of MDTs per SampleID") + theme_classic() +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), plot.margin = unit(c(1,1,1,1), "lines"))

p_coding + theme(axis.text.x = element_blank(), legend.text = element_blank()) 

svg(filename = "/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Figures/Figure_NumberOfMDTs_codinggenes.svg", width = 8, height = 6)
ggsave("/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Figures/Figure_NumberOfMDTs_codinggenes.png", plot = p, bg = "transparent", width = 6, height = 5)


print(p_coding)
dev.off()
```

```{r summaryofMDTs, warning=FALSE, message=FALSE}
summary(MDT_counts_per_samples)
MDT_counts_per_samples %>%
  dplyr::select(type,n) %>% 
  group_by(type) %>%
  summarise_all(~mean(.))

MDT_counts_per_samples_coding %>%
  dplyr::select(type,n) %>% 
  group_by(type) %>%
  summarise_all(~mean(.))

# avg of cancer samples
MDT_counts_per_samples_coding %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n)) %>% dplyr::filter(type %in% c('TCGA_pri_Mel', 'TCGA_Mel_Metastatic', 'TP_Mel', 'TCGA_Ov', 'TP_Ov', 'TP_Ov_Metastatic', 'TCGA_AML')) %>% dplyr::select(mean) %>% dplyr::summarise(mean2 = mean(mean))

# avg of gtex samples
MDT_counts_per_samples_coding %>% dplyr::group_by(type) %>% dplyr::summarise(mean = mean(n)) %>% dplyr::filter(type %in% c('GTEx_Mel', 'GTEx_Ov', 'GTEx_AML')) %>% dplyr::select(mean) %>% dplyr::summarise(mean2 = mean(mean))
```

## Read Switches
```{r read_switches, warning=FALSE, message=FALSE}
dMDTs_Ovary_tcga <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/TCGA_GTEx_ov_Switch.RDS')

dMDT_Ovary_tuPro_primary <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Pri_GTEx.RDS')

dMDT_Ovary_tuPro_metastatis <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Met_GTEx.RDS')

dMDT_Melanoma_tcga_primary <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/Melanoma/Switches_TCGA_Pri_GTEx.RDS')

dMDT_Melanoma_tcga_metastatis <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/Melanoma/Switches_TCGA_Met_GTEx.RDS')

dMDT_Melanoma_tupro <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/Melanoma/Switches_TuPro_GTEx.RDS')

dMDT_AML_tcga <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/AML/Switches_TCGA_GTEx.RDS')

dMDT_AML_tuPro <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/AML/Switches_TuPro_GTEx.RDS')
```


```{r Cohorts}
#primary
length(unique(c(dMDTs_Ovary_tcga$SampleID, dMDT_Ovary_tuPro_primary$SampleID, dMDT_Melanoma_tcga_primary$SampleID, dMDT_AML_tcga$SampleID, dMDT_AML_tuPro$SampleID)))

#metastatic
length(unique(c(dMDT_Melanoma_tupro$SampleID, dMDT_Ovary_tuPro_metastatis$SampleID)))
```


```{r makeSupplementary}
dMDT_AML_tuPro$Cohort <- 'AML_TuPro'
dMDT_AML_tcga$Cohort <- 'AML_TCGA'
dMDT_Melanoma_tupro$Cohort <- 'Melanoma_TuPro'
dMDT_Melanoma_tcga_metastatis$Cohort <- 'Melanoma_TCGA_Metastatic'
dMDT_Melanoma_tcga_primary$Cohort <- 'Melanoma_TCGA_Primary'
dMDT_Ovary_tuPro_metastatis$Cohort <- 'Ovary_TuPro_Metastatic'
dMDT_Ovary_tuPro_primary$Cohort <- 'Ovary_TuPro_Primary'
dMDTs_Ovary_tcga$Cohort <- 'Ovary_TCGA'

All_dMDTs <- rbind(dMDT_AML_tuPro, dMDT_AML_tcga, dMDT_Melanoma_tupro, dMDT_Melanoma_tcga_metastatis, dMDT_Melanoma_tcga_primary, dMDT_Ovary_tuPro_metastatis, dMDT_Ovary_tuPro_primary, dMDTs_Ovary_tcga)

colnames(All_dMDTs) <- c('SampleID', 'ENSG', 'dMDT', 'ENST2', 'TPM1', 'TPM2', 'enrichment', 'p_val', 'relative_cancer_exp', 'rel_GTEX_exp', "MDT_in_GTEx", 'adj_p_value', "Cohort")

#write.table(All_dMDTs, "/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Tables/dMDTs_all.tsv")
#write.table(All_dMDTs, "/Users/tulaykarakulak/Documents/PhD/Projects/TumorProfiler/Analysis/May2024/Tables/Supplementaries/SupplementaryTable1.tsv")
```

length(unique(All_dMDTs$dMDT))
[1] 6114 unique dMDTs

 length(unique(All_dMDTs$ENSG))
[1] 4151 unique genes

TP53: 10 Ovary TCGA and 1 TuPro Primary and 9 TuPro Metastatic

## Figure 2C: Number of cMDTs across Samples
```{r statistics_switch_analysis, warning=FALSE, message=FALSE}
library(dplyr)
Switches_in_Dataset_gtex_tcga_Ov_n_per_sample <- dMDTs_Ovary_tcga %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tp_Ov_n_per_sample <- dMDT_Ovary_tuPro_primary %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tp_metas_Ov_n_per_sample <- dMDT_Ovary_tuPro_metastatis %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tcga_Mel_n_per_sample <- dMDT_Melanoma_tcga_primary %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tcga_metas_Mel_n_per_sample <- dMDT_Melanoma_tcga_metastatis %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tp_Mel_n_per_sample <- dMDT_Melanoma_tupro %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tcga_AML_n_per_sample <- dMDT_AML_tcga %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 

Switches_in_Dataset_gtex_tp_AML_n_per_sample <- dMDT_AML_tuPro %>% dplyr::group_by(SampleID) %>% dplyr::summarise(n=n(), across()) %>% arrange(desc(n)) %>% dplyr::select(SampleID, n) %>% dplyr::distinct() 


dMDT_counts_per_samples <- rbind(cbind(type = rep('TCGA_Mel', dim(Switches_in_Dataset_gtex_tcga_Mel_n_per_sample)[1]), Switches_in_Dataset_gtex_tcga_Mel_n_per_sample[,1:2]),
                                   cbind(type = rep('TCGA_Mel_Metastatic', dim(Switches_in_Dataset_gtex_tcga_metas_Mel_n_per_sample)[1]), Switches_in_Dataset_gtex_tcga_metas_Mel_n_per_sample[,1:2]),
                        cbind(type = rep('TP_Mel', dim(Switches_in_Dataset_gtex_tp_Mel_n_per_sample)[1]), Switches_in_Dataset_gtex_tp_Mel_n_per_sample[,1:2]),
                        cbind(type = rep('TCGA_Ov', dim(Switches_in_Dataset_gtex_tcga_Ov_n_per_sample)[1]), Switches_in_Dataset_gtex_tcga_Ov_n_per_sample[,1:2]), 
                        cbind(type = rep('TP_Ov', dim(Switches_in_Dataset_gtex_tp_Ov_n_per_sample)[1]), Switches_in_Dataset_gtex_tp_Ov_n_per_sample[,1:2]), 
                        cbind(type = rep('TP_Ov_Metastatic', dim(Switches_in_Dataset_gtex_tp_metas_Ov_n_per_sample)[1]), Switches_in_Dataset_gtex_tp_metas_Ov_n_per_sample[,1:2]),
                        cbind(type = rep('TCGA_AML', dim(Switches_in_Dataset_gtex_tcga_AML_n_per_sample)[1]), Switches_in_Dataset_gtex_tcga_AML_n_per_sample[,1:2]),cbind(type = rep('TP_AML', dim(Switches_in_Dataset_gtex_tp_AML_n_per_sample)[1]), Switches_in_Dataset_gtex_tp_AML_n_per_sample[,1:2]) )

library(ggplot2)
library(RColorBrewer)

dMDT_counts_per_samples$type <- factor(dMDT_counts_per_samples$type, levels = unique(dMDT_counts_per_samples$type))

# If you have 3 main types, you can choose a palette of 3 colors.
my_palette <- c("#d85bd4",
"#db95d9",
"#dcb7dc",
"#2d858d",
"#34cacd",
"#73e5ea",
"#df8250",
"#e4a481")

means_df_dMDT <- dMDT_counts_per_samples %>%
  group_by(type) %>%
  summarise(mean_value = mean(n))

# Find the maximum y value to place the text above everything else
max_y_dMDT <- max(dMDT_counts_per_samples$n, na.rm = TRUE)

a <- ggplot(dMDT_counts_per_samples, aes(x = type, y = n, color = type)) +
    geom_boxplot() +
    geom_text(data = means_df_dMDT, aes(x = type, y = max_y_dMDT+10, label = round(mean_value, 1)), 
            position = position_nudge(y = 500), vjust = 0, hjust = 1, size = 3, color = "darkgray", angle=90) +
    scale_color_manual(values = my_palette) + # use the color palette 
    xlab("Cohort") + 
    ylab("Number of dMDTs") +
    ggtitle("Distribution of Number of dMDTs per SampleID") +
    theme_classic() + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


a
```

```{r summaryofdMDTs, warning=FALSE, message=FALSE}
summary(dMDT_counts_per_samples)
dMDT_counts_per_samples %>%
  dplyr::select(type,n) %>% 
  group_by(type) %>%
  summarise_all(~mean(.))
```


```{r removeSingleratons}
calc_singletons <- function (cMDT_file) {
  
  cMDT_file %>% group_by(ENSG, dMDT) %>%
    summarise(
      n_samples = n_distinct(SampleID), 
      .groups = 'drop'
    )
}

dMDT_AML <- rbind(dMDT_AML_tuPro, dMDT_AML_tcga)
dMDT_AML_sing <- calc_singletons(dMDT_AML)

dMDT_Melanoma <- rbind(dMDT_Melanoma_tupro, dMDT_Melanoma_tcga_metastatis, dMDT_Melanoma_tcga_primary)
dMDT_Melanoma_sing <- calc_singletons(dMDT_Melanoma)

dMDT_Ovary <- rbind(dMDT_Ovary_tuPro_metastatis, dMDT_Ovary_tuPro_primary, dMDTs_Ovary_tcga)
dMDT_Ovary_sing <- calc_singletons(dMDT_Ovary)


dMDT_AML_sing_1 <- unique(dMDT_AML_sing[dMDT_AML_sing$n_samples == 1, 'dMDT'])$dMDT
dMDT_Melanoma_sing_1 <- unique(dMDT_Melanoma_sing[dMDT_Melanoma_sing$n_samples == 1, 'dMDT'])$dMDT
dMDT_Ovary_sing_1 <- unique(dMDT_Ovary_sing[dMDT_Ovary_sing$n_samples == 1, 'dMDT'])$dMDT
```


## Figure 2D: Check Overlapping cMDTs among Cohorts

```{r checkOverlaps}
##library(UpSetR)

dMDT_AML_tuPro$Cohort <- 'AML_TuPro'
dMDT_AML_tcga$Cohort <- 'AML_TCGA'
dMDT_Melanoma_tupro$Cohort <- 'Melanoma_TuPro'
dMDT_Melanoma_tcga_metastatis$Cohort <- 'Melanoma_TCGA_Metastatic'
dMDT_Melanoma_tcga_primary$Cohort <- 'Melanoma_TCGA_Primary'
dMDT_Ovary_tuPro_metastatis$Cohort <- 'Ovary_TuPro_Metastatic'
dMDT_Ovary_tuPro_primary$Cohort <- 'Ovary_TuPro_Primary'
dMDTs_Ovary_tcga$Cohort <- 'Ovary_TCGA'

##  filter cMDTs seen in a single sample
dMDT_AML_tuPro_filtered <- dMDT_AML_tuPro[!dMDT_AML_tuPro$dMDT %in% dMDT_AML_sing_1,]
dMDT_AML_tcga_filtered <- dMDT_AML_tcga[!dMDT_AML_tcga$dMDT %in% dMDT_AML_sing_1,]

dMDT_Melanoma_tupro_filtered <- dMDT_Melanoma_tupro[!dMDT_Melanoma_tupro$dMDT %in% dMDT_Melanoma_sing_1,]
dMDT_Melanoma_tcga_metastatis_filtered <- dMDT_Melanoma_tcga_metastatis[!dMDT_Melanoma_tcga_metastatis$dMDT %in% dMDT_Melanoma_sing_1,]
dMDT_Melanoma_tcga_primary_filtered <- dMDT_Melanoma_tcga_primary[!dMDT_Melanoma_tcga_primary$dMDT %in% dMDT_Melanoma_sing_1,]

dMDT_Ovary_tuPro_metastatis_filtered <- dMDT_Ovary_tuPro_metastatis[!dMDT_Ovary_tuPro_metastatis$dMDT %in% dMDT_Ovary_sing_1,]
dMDT_Ovary_tuPro_primary_filtered <- dMDT_Ovary_tuPro_primary[!dMDT_Ovary_tuPro_primary$dMDT %in% dMDT_Ovary_sing_1,]
dMDTs_Ovary_tcga_filtered <- dMDTs_Ovary_tcga[!dMDTs_Ovary_tcga$dMDT %in% dMDT_Ovary_sing_1,]


All_dMDT_list <- list(TuPro_AML = unique(dMDT_AML_tuPro_filtered$dMDT), TCGA_AML = unique(dMDT_AML_tcga_filtered$dMDT),
                      TuPro_Mel = unique(dMDT_Melanoma_tupro_filtered$dMDT), 
                      TCGA_Mel_Metastatic = unique(dMDT_Melanoma_tcga_metastatis_filtered$dMDT), 
                      TCGA_Mel_Pri = unique(dMDT_Melanoma_tcga_primary_filtered$dMDT), 
                      TuPro_Ovary_Metastatic = unique(dMDT_Ovary_tuPro_metastatis_filtered$dMDT),
                      TuPro_Ovary_Primary = unique(dMDT_Ovary_tuPro_primary_filtered$dMDT), 
                      TCGA_Ovary = unique(dMDTs_Ovary_tcga_filtered$dMDT))

##upset(fromList(All_dMDT_list), order.by = 'freq', nsets = 8)

##All_overlap <- ggVennDiagram::ggVennDiagram(All_dMDT_list) # output is upset plot
##All_overlap

## ggVenn Ovary
All_dMDT_Ovary <- list(TuPro_Ovary_Metastatic = unique(dMDT_Ovary_tuPro_metastatis_filtered$dMDT),
                      TuPro_Ovary_Primary = unique(dMDT_Ovary_tuPro_primary_filtered$dMDT), 
                      TCGA_Ovary = unique(dMDTs_Ovary_tcga_filtered$dMDT))
Ovary_overlap <- ggVennDiagram::ggVennDiagram(All_dMDT_Ovary, , label = "none")   +  scale_fill_gradient(low="grey90",high = "turquoise")
Ovary_overlap
# ggVenn Melanoma
All_dMDT_Mel <- list(TuPro_Mel = unique(dMDT_Melanoma_tupro_filtered$dMDT), 
                      TCGA_Mel_Metastatic = unique(dMDT_Melanoma_tcga_metastatis_filtered$dMDT), 
                      TCGA_Mel_Pri = unique(dMDT_Melanoma_tcga_primary_filtered$dMDT))
Mel_overlap <- ggVennDiagram::ggVennDiagram(All_dMDT_Mel, , label = "none") +  scale_fill_gradient(low="grey90",high = "purple")
Mel_overlap

# ggVenn AML
All_dMDT_AML <- list(TuPro_AML = unique(dMDT_AML_tuPro_filtered$dMDT), TCGA_AML = unique(dMDT_AML_tcga_filtered$dMDT))
AML_overlap <- ggVennDiagram::ggVennDiagram(All_dMDT_AML, category.names = c("", ""), label = "none")  +  scale_fill_gradient(low="grey90",high = "orange")
AML_overlap 

```


## Mostly Found dMDTs across cohorts
```{r FrequencyOfdMDTs}

dMDT_AML_tuPro$Cancer <- 'AML'
dMDT_AML_tcga$Cancer <- 'AML'
dMDT_Melanoma_tupro$Cancer <- 'Melanoma'
dMDT_Melanoma_tcga_metastatis$Cancer <- 'Melanoma'
dMDT_Melanoma_tcga_primary$Cancer <- 'Melanoma'
dMDT_Ovary_tuPro_metastatis$Cancer <- 'Ovary'
dMDT_Ovary_tuPro_primary$Cancer <- 'Ovary'
dMDTs_Ovary_tcga$Cancer <- 'Ovary'

All_dMDTs2 <- rbind(dMDT_AML_tuPro, dMDT_AML_tcga, dMDT_Melanoma_tupro, dMDT_Melanoma_tcga_metastatis, dMDT_Melanoma_tcga_primary, dMDT_Ovary_tuPro_metastatis, dMDT_Ovary_tuPro_primary, dMDTs_Ovary_tcga)

calc_singletons2 <- function (cMDT_file) {
  
  cMDT_file %>% group_by(ENSG, dMDT, Cancer) %>%
    summarise(
      n_samples = n_distinct(SampleID), 
      .groups = 'drop'
    )
}

All_dMDTs2_singletons <- calc_singletons2(All_dMDTs2)
length(unique(All_dMDTs2_singletons[All_dMDTs2_singletons$n_samples == 1,'dMDT'])$dMDT)/length(unique(All_dMDTs2$dMDT))*100 #percentage of dMDTs found in one sample - 41.00 % dMDTs are singletons when we filter by a Cancer Type. 


dMDT_counts <- All_dMDTs %>% group_by(ENSG, dMDT, Cohort) %>% dplyr::summarise(count = n()) %>% dplyr::arrange(desc(count))

Unique_samples <- All_dMDTs %>% group_by(Cohort) %>%
  summarise(total_samples = n_distinct(SampleID), .groups = 'drop')

percentage_data <- merge(dMDT_counts, Unique_samples, by = "Cohort") %>% mutate(percentage = (count / total_samples) * 100) %>%
  arrange(desc(count))
percentage_data$percentage <- format(round(percentage_data$percentage, 2), nsmall = 2)
head(percentage_data)
# write.table(percentage_data, "SupplementaryTable2.tsv")
```


```{r morethan30Percent}
percentage_data_30perc <- percentage_data %>% dplyr::filter(percentage >= 30)

dim(percentage_data_30perc)
length(unique(percentage_data_30perc$dMDT))
```

# check frequently found dMDTs in both cohorts - overlapping dMDTs in both cohorts
```{r frequentandcommondMDTs}
Ovary_TCGA_30 <- percentage_data[percentage_data$Cohort == 'Ovary_TCGA', ] %>% dplyr::filter(percentage >= 30)
Ovary_TuPro_Metastatic_30 <- percentage_data[percentage_data$Cohort == 'Ovary_TuPro_Metastatic', ] %>% dplyr::filter(percentage >= 30)
Ovary_TuPro_Primary_30 <- percentage_data[percentage_data$Cohort == 'Ovary_TuPro_Primary', ] %>% dplyr::filter(percentage >= 30)
Melanoma_TuPro_30 <- percentage_data[percentage_data$Cohort == 'Melanoma_TuPro', ] %>% dplyr::filter(percentage >= 30)
Melanoma_TCGA_Metastatic_30 <- percentage_data[percentage_data$Cohort == 'Melanoma_TCGA_Metastatic', ] %>% dplyr::filter(percentage >= 30)
Melanoma_TCGA_Primary_30 <- percentage_data[percentage_data$Cohort == 'Melanoma_TCGA_Primary', ] %>% dplyr::filter(percentage >= 30)
AML_TuPro_30 <- percentage_data[percentage_data$Cohort == 'AML_TuPro', ] %>% dplyr::filter(percentage >= 30)
AML_TCGA_30 <- percentage_data[percentage_data$Cohort == 'AML_TCGA', ] %>% dplyr::filter(percentage >= 30)

All_dMDT_AML_vector <- intersect(AML_TuPro_30$dMDT, AML_TCGA_30$dMDT)
All_dMDT_Mel_met_vector <- intersect(Melanoma_TCGA_Metastatic_30$dMDT, Melanoma_TuPro_30$dMDT)
All_dMDT_Mel_pri_vector <- intersect(Melanoma_TCGA_Primary_30$dMDT, Melanoma_TuPro_30$dMDT)
All_dMDT_Ov_met_vector <- intersect(Ovary_TCGA_30$dMDT, Ovary_TuPro_Metastatic_30$dMDT)
All_dMDT_Ov_pri_vector <- intersect(Ovary_TCGA_30$dMDT, Ovary_TuPro_Primary_30$dMDT)


Ovary_TCGA_10 <- percentage_data[percentage_data$Cohort == 'Ovary_TCGA', ] %>% dplyr::filter(percentage >= 10)
Ovary_TuPro_Metastatic_10 <- percentage_data[percentage_data$Cohort == 'Ovary_TuPro_Metastatic', ] %>% dplyr::filter(percentage >= 10)
Ovary_TuPro_Primary_10 <- percentage_data[percentage_data$Cohort == 'Ovary_TuPro_Primary', ] %>% dplyr::filter(percentage >= 10)

Melanoma_TuPro_10 <- percentage_data[percentage_data$Cohort == 'Melanoma_TuPro', ] %>% dplyr::filter(percentage >= 10)
Melanoma_TCGA_Metastatic_10 <- percentage_data[percentage_data$Cohort == 'Melanoma_TCGA_Metastatic', ] %>% dplyr::filter(percentage >= 10)
Melanoma_TCGA_Primary_10 <- percentage_data[percentage_data$Cohort == 'Melanoma_TCGA_Primary', ] %>% dplyr::filter(percentage >= 10)

All_dMDT_Ov_met_vector_10 <- intersect(Ovary_TCGA_10$dMDT, Ovary_TuPro_Metastatic_10$dMDT)
All_dMDT_Ov_pri_vector_10 <- intersect(Ovary_TCGA_10$dMDT, Ovary_TuPro_Primary_10$dMDT)
All_dMDT_Mel_met_vector_10 <- intersect(Melanoma_TCGA_Metastatic_10$dMDT, Melanoma_TuPro_10$dMDT)
All_dMDT_Mel_pri_vector_10 <- intersect(Melanoma_TCGA_Primary_10$dMDT, Melanoma_TuPro_10$dMDT)
```


```{r percentagesOfCommons}
AML_Biomarker_candidates <- percentage_data %>% dplyr::filter(percentage >= 30) %>% dplyr::filter(Cohort %in% c('AML_TuPro', 'AML_TCGA')) %>% dplyr::filter(dMDT %in% All_dMDT_AML_vector)

Ov_Biomarker_candidates_met  <- percentage_data  %>% dplyr::filter(percentage >= 30)  %>% dplyr::filter(Cohort %in% c('Ovary_TCGA', 'Ovary_TuPro_Metastatic')) %>% dplyr::filter(dMDT %in% c(All_dMDT_Ov_met_vector))

Ov_Biomarker_candidates_pri <- percentage_data  %>% dplyr::filter(percentage >= 30)  %>% dplyr::filter(Cohort %in% c('Ovary_TCGA', 'Ovary_TuPro_Primary')) %>% dplyr::filter(dMDT %in% c(All_dMDT_Ov_pri_vector))

Mel_Biomarker_candidates_met <- percentage_data %>% dplyr::filter(percentage >= 30) %>% dplyr::filter(Cohort %in% c('Melanoma_TCGA_Metastatic', 'Melanoma_TuPro')) %>% dplyr::filter(dMDT %in% c(All_dMDT_Mel_met_vector))

Mel_Biomarker_candidates_pri <- percentage_data %>% dplyr::filter(percentage >= 30) %>% dplyr::filter(Cohort %in% c('Melanoma_TCGA_Primary', 'Melanoma_TuPro')) %>% dplyr::filter(dMDT %in% c(All_dMDT_Mel_pri_vector))


All_Biomarker_candidates <- rbind(AML_Biomarker_candidates, Ov_Biomarker_candidates_met, Ov_Biomarker_candidates_pri, Mel_Biomarker_candidates_met, Mel_Biomarker_candidates_pri)

All_Biomarker_candidates <- All_Biomarker_candidates %>% dplyr::distinct()
```


# Check Transmembrane Proteins
```{r readIDs}
TM_proteins <- read.csv('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Transmembrane_Proteins/idmapping_2024_05_11.tsv', header = TRUE, sep = '\t')
TM_proteins$To <- gsub("\\..*", "", TM_proteins$To)
head(TM_proteins)
```

```{r TMindMDT}
All_dMDTs_TM <- All_dMDTs[All_dMDTs$ENSG %in% TM_proteins$To,]
dim(All_dMDTs_TM)
head(All_dMDTs_TM)
length(unique(All_dMDTs_TM$dMDT)) # 1707 dMDT encoding transmembrane proteins
All_dMDTs_TM_percentage <- percentage_data %>% dplyr::filter(percentage >= 30) %>% dplyr::filter(dMDT %in% All_dMDTs_TM$dMDT)

head(All_dMDTs_TM_percentage)
dim(All_dMDTs_TM_percentage)
```

