---
title: "Manuscript_11"
output: html_document
date: "2025-11-12"
---

Transcript and CDS Length of Transcripts/cMDTs - Supplementary Figure6 (cMDT specific part only)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#required libraries
```{r loadlibraries, echo=FALSE}
library(tidyverse)
library(tidyr)
library(dplyr)
library(ggplot2)
library(pheatmap)
library(sva)
library(stats)
library(data.table)
library(BSDA)
library(Biostrings)
library(ggvenn)
```


# Melanoma Samples
```{r read_output_diann}
library(readr)
Main_output <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/Melanoma/Main_Report.tsv', delim = '\t', col_names = TRUE)
dim(Main_output)
Main_output <- Main_output %>% filter(!str_detect(Genes, ";")) 
Main_output <- Main_output[Main_output$Q.Value <= 0.01 & Main_output$PG.Q.Value <= 0.01, ]
head(Main_output)
dim(Main_output)
```

```{r read_Library}
Library_Mel <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/ProteomicsLibrary/Melanoma/ensg_enst_ensp_seq_proteomics_ready.tsv', delim = ',', col_names = TRUE)
head(Library_Mel)
```

We consider all Melanoma Library - nothing specific to dMDTs! 
```{r isolateProteins}
ENSP_w_Unique_Peptides <- unique(Main_output[Main_output$Protein.Ids %in% Library_Mel$Protein.stable.ID, 'Protein.Ids']$Protein.Ids)
ENSP_wo_unique_Peptides <- unique(Library_Mel$Protein.stable.ID[!Library_Mel$Protein.stable.ID %in% ENSP_w_Unique_Peptides])
```


## Transcript Length - which includes UTRs and CDS 
```{r ReadBioMart}
ENSG_ENST_ENSP_Length <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Biomart/ENSG_ENST_ENSP_Length.txt', delim = '\t', col_names = TRUE)
head(ENSG_ENST_ENSP_Length)
```


```{r MergeFiles}
ENST_Length_w_Peptides <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_w_Unique_Peptides,]
ENST_Length_wo_Peptides <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_wo_unique_Peptides,]
head(ENST_Length_w_Peptides)
colnames(ENST_Length_w_Peptides) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_Length_wo_Peptides) <- c('ENSG', 'ENST', 'ENSP', 'Length')
```
mean(ENST_Length_w_Peptides$Length)
[1] 3259
mean(ENST_Length_wo_Peptides$Length)
[1] 2000


```{r graphs}
coul2 <- c("lightblue", "gray")
ENST_Length_w_Peptides$Evidence <- 'PeptideEvidence'
ENST_Length_wo_Peptides$Evidence <- 'NoPeptideEvidence'
ENST_Lengths_PepEvidence <- rbind(ENST_Length_w_Peptides, ENST_Length_wo_Peptides)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot <- ggplot(ENST_Lengths_PepEvidence , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot <- ggplot(ENST_Lengths_PepEvidence , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

library(cowplot)
Melanoma_dMDT_Library_TranscriptLength <- plot_grid(boxplot, density_plot, align = "v", ncol = 1, rel_heights = c(1, 5))
Melanoma_dMDT_Library_TranscriptLength

svg(filename = "/Users/admin/Desktop/Melanoma_dMDT_Library_TranscriptLength.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Melanoma_dMDT_Library_TranscriptLength.png", plot = Melanoma_dMDT_Library_TranscriptLength, bg = "transparent", width = 6, height = 5)
print(Melanoma_dMDT_Library_TranscriptLength)
dev.off()
```

wilcox.test(ENST_Lengths_PepEvidence[ENST_Lengths_PepEvidence$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_Lengths_PepEvidence[ENST_Lengths_PepEvidence$Evidence == 'NoPeptideEvidence', 'Length']$Length)

Wilcoxon rank sum test with continuity correction

data:  ENST_Lengths_PepEvidence[ENST_Lengths_PepEvidence$Evidence == "PeptideEvidence", "Length"]$Length and ENST_Lengths_PepEvidence[ENST_Lengths_PepEvidence$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 9e+06, p-value <2e-16
alternative hypothesis: true location shift is not equal to 0

## CDS Length only 
Maybe it would make a difference 
```{r read_CDSLength}
ENSG_ENST_ENSP_CDSLength <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Biomart/Ensembl105//ENSG_ENST_ENSP_CDS_Length.txt', delim = '\t', col_names = TRUE)
head(ENSG_ENST_ENSP_CDSLength)

ENST_CDSLength_w_Peptides <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_w_Unique_Peptides,]
ENST_CDSLength_wo_Peptides <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_wo_unique_Peptides,]
head(ENST_CDSLength_w_Peptides)
colnames(ENST_CDSLength_w_Peptides) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_CDSLength_wo_Peptides) <- c('ENSG', 'ENST', 'ENSP', 'Length')


ENST_CDSLength_w_Peptides$Evidence <- 'PeptideEvidence'
ENST_CDSLength_wo_Peptides$Evidence <- 'NoPeptideEvidence'

ENST_CDSLengths_PepEvidence <- rbind(ENST_CDSLength_w_Peptides, ENST_CDSLength_wo_Peptides)

density_plot_CDS <- ggplot(ENST_CDSLengths_PepEvidence , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_CDS <- ggplot(ENST_CDSLengths_PepEvidence , aes(y =log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

Melanoma_dMDT_Library_CDSLength <- plot_grid(boxplot_CDS, density_plot_CDS, align = "v", ncol = 1, rel_heights = c(1, 5))

Melanoma_dMDT_Library_CDSLength

svg(filename = "/Users/admin/Desktop/Melanoma_dMDT_Library_CDSLength.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Melanoma_dMDT_Library_CDSLength.png", plot = Melanoma_dMDT_Library_CDSLength, bg = "transparent", width = 6, height = 5)
print(Melanoma_dMDT_Library_CDSLength)
dev.off()
```
 wilcox.test(ENST_CDSLengths_PepEvidence[ENST_CDSLengths_PepEvidence$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_CDSLengths_PepEvidence[ENST_CDSLengths_PepEvidence$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

data:  ENST_CDSLengths_PepEvidence[ENST_CDSLengths_PepEvidence$Evidence == "PeptideEvidence", "Length"]$Length and ENST_CDSLengths_PepEvidence[ENST_CDSLengths_PepEvidence$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 9e+06, p-value <2e-16
alternative hypothesis: true location shift is not equal to 0

## Ovarian Samples
```{r read_output_diann}
library(readr)
Main_output_Ovary <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/Ovary/DIANN/Main_Report.tsv', delim = '\t', col_names = TRUE)
Main_output_Ovary <- Main_output_Ovary %>% filter(!str_detect(Genes, ";"))
Main_output_Ovary <- Main_output_Ovary[Main_output_Ovary$Q.Value <= 0.01 & Main_output_Ovary$PG.Q.Value <= 0.01, ]
```


```{r read_Library}
Library_Ov <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/ProteomicsLibrary/Ovary/ensg_enst_ensp_seq_proteomics_ready.tsv', delim = ',', col_names = TRUE)
head(Library_Ov)

ENSP_w_Unique_Peptides_Ov <- unique(Main_output_Ovary[Main_output_Ovary$Protein.Ids %in% Library_Ov$Protein.stable.ID, 'Protein.Ids']$Protein.Ids)
ENSP_wo_unique_Peptides_Ov <- unique(Library_Ov$Protein.stable.ID[!Library_Ov$Protein.stable.ID %in% ENSP_w_Unique_Peptides_Ov])

ENST_Length_w_Peptides_Ov <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_w_Unique_Peptides_Ov,]
ENST_Length_wo_Peptides_Ov <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_wo_unique_Peptides_Ov,]
head(ENST_Length_w_Peptides_Ov)
colnames(ENST_Length_w_Peptides_Ov) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_Length_wo_Peptides_Ov) <- c('ENSG', 'ENST', 'ENSP', 'Length')


ENST_Length_w_Peptides_Ov$Evidence <- 'PeptideEvidence'
ENST_Length_wo_Peptides_Ov$Evidence <- 'NoPeptideEvidence'

ENST_Lengths_PepEvidence_Ov <- rbind(ENST_Length_w_Peptides_Ov, ENST_Length_wo_Peptides_Ov)

Ov_Density_plot <- ggplot(ENST_Lengths_PepEvidence_Ov , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

Ov_boxplot <- ggplot(ENST_Lengths_PepEvidence_Ov , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())


Ov_dMDT_Library_TranscriptLength <- plot_grid(Ov_boxplot, Ov_Density_plot, align = "v", ncol = 1, rel_heights = c(1, 5))

Ov_dMDT_Library_TranscriptLength

svg(filename = "/Users/admin/Desktop/Ov_dMDT_Library_TranscriptLength.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Ov_dMDT_Library_TranscriptLength.png", plot = Ov_dMDT_Library_TranscriptLength, bg = "transparent", width = 6, height = 5)
print(Ov_dMDT_Library_TranscriptLength)
dev.off()
```

# Ovarian CDS Length
```{r CDSLength}
ENST_CDSLength_w_Peptides_ov <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_w_Unique_Peptides_Ov,]
ENST_CDSLength_wo_Peptides_ov <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_wo_unique_Peptides_Ov,]
head(ENST_CDSLength_w_Peptides)

colnames(ENST_CDSLength_w_Peptides_ov) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_CDSLength_wo_Peptides_ov) <- c('ENSG', 'ENST', 'ENSP', 'Length')


ENST_CDSLength_w_Peptides_ov$Evidence <- 'PeptideEvidence'
ENST_CDSLength_wo_Peptides_ov$Evidence <- 'NoPeptideEvidence'

ENST_CDSLengths_PepEvidence_Ov <- rbind(ENST_CDSLength_w_Peptides_ov, ENST_CDSLength_wo_Peptides_ov)

Ov_Density_plot_CDS <- ggplot(ENST_CDSLengths_PepEvidence_Ov , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

Ov_boxplot_CDS <- ggplot(ENST_CDSLengths_PepEvidence_Ov , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())


Ov_dMDT_Library_CDSLength <- plot_grid(Ov_boxplot_CDS, Ov_Density_plot_CDS, align = "v", ncol = 1, rel_heights = c(1, 5))

Ov_dMDT_Library_CDSLength

svg(filename = "/Users/admin/Desktop/Ov_dMDT_Library_CDSLength.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Ov_dMDT_Library_CDSLength.png", plot = Ov_dMDT_Library_CDSLength, bg = "transparent", width = 6, height = 5)
print(Ov_dMDT_Library_CDSLength)
dev.off()
```

wilcox.test(ENST_Lengths_PepEvidence_Ov[ENST_Lengths_PepEvidence_Ov$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_Lengths_PepEvidence_Ov[ENST_Lengths_PepEvidence_Ov$Evidence == 'NoPeptideEvidence', 'Length']$Length)

Wilcoxon rank sum test with continuity correction

data:  ENST_Lengths_PepEvidence_Ov[ENST_Lengths_PepEvidence_Ov$Evidence == "PeptideEvidence", "Length"]$Length and ENST_Lengths_PepEvidence_Ov[ENST_Lengths_PepEvidence_Ov$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 5e+06, p-value <2e-16
alternative hypothesis: true location shift is not equal to 0


## AML Dataset 
```{r checkAMLresults}
library(readr)
Main_output_AML <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/AML/Main_Report.tsv', delim = '\t', col_names = TRUE)
Main_output_AML <- Main_output_AML %>% filter(!str_detect(Genes, ";"))
Main_output_AML <- Main_output_AML[Main_output_AML$Q.Value <= 0.01 & Main_output_AML$PG.Q.Value <= 0.01, ]

Library_AML <- read_delim('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Proteomics/ProteomicsLibrary/AML/ensg_enst_ensp_seq_proteomics_ready.tsv', delim = ',', col_names = TRUE)

ENSP_w_Unique_Peptides_AML <- unique(Main_output_AML[Main_output_AML$Protein.Ids %in% Library_AML$Protein.stable.ID, 'Protein.Ids']$Protein.Ids)

ENSP_wo_unique_Peptides_AML <- unique(Library_AML$Protein.stable.ID[!Library_AML$Protein.stable.ID %in% ENSP_w_Unique_Peptides_AML])


ENST_Length_w_Peptides_AML <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_w_Unique_Peptides_AML,]
ENST_Length_wo_Peptides_AML <- ENSG_ENST_ENSP_Length[ENSG_ENST_ENSP_Length$`Protein stable ID` %in% ENSP_wo_unique_Peptides_AML,]

head(ENST_Length_w_Peptides_AML)
colnames(ENST_Length_w_Peptides_AML) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_Length_wo_Peptides_AML) <- c('ENSG', 'ENST', 'ENSP', 'Length')


ENST_Length_w_Peptides_AML$Evidence <- 'PeptideEvidence'
ENST_Length_wo_Peptides_AML$Evidence <- 'NoPeptideEvidence'

ENST_Lengths_PepEvidence_AML <- rbind(ENST_Length_w_Peptides_AML, ENST_Length_wo_Peptides_AML)

ggplot(ENST_Lengths_PepEvidence_AML , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

ggplot(ENST_Lengths_PepEvidence_AML , aes(x=log10(Length), color=Evidence)) +
  geom_boxplot() +
  scale_color_manual(values=coul2) + theme_classic()

## CDS Length
ENST_CDSLength_w_Peptides_AML <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_w_Unique_Peptides_AML,]

ENST_CDSLength_wo_Peptides_AML <- ENSG_ENST_ENSP_CDSLength[ENSG_ENST_ENSP_CDSLength$`Protein stable ID` %in% ENSP_wo_unique_Peptides_AML,]
head(ENST_CDSLength_w_Peptides)


colnames(ENST_CDSLength_w_Peptides_AML) <- c('ENSG', 'ENST', 'ENSP', 'Length')
colnames(ENST_CDSLength_wo_Peptides_AML) <- c('ENSG', 'ENST', 'ENSP', 'Length')


ENST_CDSLength_w_Peptides_AML$Evidence <- 'PeptideEvidence'
ENST_CDSLength_wo_Peptides_AML$Evidence <- 'NoPeptideEvidence'

ENST_CDSLengths_PepEvidence_AML <- rbind(ENST_CDSLength_w_Peptides_AML, ENST_CDSLength_wo_Peptides_AML)

AML_Density_plot_CDS <- ggplot(ENST_CDSLengths_PepEvidence_AML , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

AML_boxplot_CDS <- ggplot(ENST_CDSLengths_PepEvidence_AML , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())


AML_dMDT_Library_CDSLength <- plot_grid(AML_boxplot_CDS, AML_Density_plot_CDS, align = "v", ncol = 1, rel_heights = c(1, 5))

AML_dMDT_Library_CDSLength

svg(filename = "/Users/admin/Desktop/AML_dMDT_Library_CDSLength.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/AML_dMDT_Library_CDSLength.png", plot = AML_dMDT_Library_CDSLength, bg = "transparent", width = 6, height = 5)
print(AML_dMDT_Library_CDSLength)
dev.off()
```


## cMDT specific Analysis
```{r Switches}
Switches_Melanoma <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/Melanoma/Switches_TuPro_GTEx.RDS')
ensg_enst_ensp_apris_v104 <-  read.csv('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/Biomart/mart_export-2May2023.txt', sep='\t')
head(ensg_enst_ensp_apris_v104)
Switches_Melanoma_ensps <- merge(ensg_enst_ensp_apris_v104, Switches_Melanoma, by.x= 'Transcript.stable.ID', by.y = 'dMDT', sort=FALSE)
```

## dMDT Level Analysis Melanoma
```{r IdentifiedvsNonIdentifieddMDTs}
ENST_Length_w_Peptides_dMDT <- ENST_Length_w_Peptides[ENST_Length_w_Peptides$ENSP %in% Switches_Melanoma_ensps$Protein.stable.ID,]
ENST_Length_wo_Peptides_dMDT <- ENST_Length_wo_Peptides[ENST_Length_wo_Peptides$ENSP %in% Switches_Melanoma_ensps$Protein.stable.ID,]

ENST_Lengths_PepEvidence_dMDT <- rbind(ENST_Length_w_Peptides_dMDT, ENST_Length_wo_Peptides_dMDT)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot_dMDT <- ggplot(ENST_Lengths_PepEvidence_dMDT , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT <- ggplot(ENST_Lengths_PepEvidence_dMDT , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

library(cowplot)
Melanoma_dMDT_Library_TranscriptLength_dMDT <- plot_grid(boxplot_dMDT, density_plot_dMDT, align = "v", ncol = 1, rel_heights = c(1, 5))

Melanoma_dMDT_Library_TranscriptLength_dMDT


svg(filename = "/Users/admin/Desktop/Melanoma_dMDT_Library_TranscriptLength_cMDT.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Melanoma_dMDT_Library_TranscriptLength_dMDT.png", plot = Melanoma_dMDT_Library_TranscriptLength_dMDT, bg = "transparent", width = 6, height = 5)
print(Melanoma_dMDT_Library_TranscriptLength_dMDT)
dev.off()

## CDS Length 
ENST_CDSLength_w_Peptides_dMDT <- ENST_CDSLength_w_Peptides[ENST_CDSLength_w_Peptides$ENSP %in% Switches_Melanoma_ensps$Protein.stable.ID,]
ENST_CDSLength_wo_Peptides_dMDT <- ENST_CDSLength_wo_Peptides[ENST_CDSLength_wo_Peptides$ENSP %in% Switches_Melanoma_ensps$Protein.stable.ID,]

ENST_CDSLengths_PepEvidence_dMDT <- rbind(ENST_CDSLength_w_Peptides_dMDT, ENST_CDSLength_wo_Peptides_dMDT)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot_dMDT_CDS <- ggplot(ENST_CDSLengths_PepEvidence_dMDT , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT_CDS <- ggplot(ENST_CDSLengths_PepEvidence_dMDT , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

library(cowplot)
Melanoma_dMDT_Library_CDSLength_dMDT <- plot_grid(boxplot_dMDT_CDS, density_plot_dMDT_CDS, align = "v", ncol = 1, rel_heights = c(1, 5))

Melanoma_dMDT_Library_CDSLength_dMDT


svg(filename = "/Users/admin/Desktop/Melanoma_dMDT_Library_CDSLength_dMDT.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Melanoma_dMDT_Library_CDSLength_dMDT.png", plot = Melanoma_dMDT_Library_CDSLength_dMDT, bg = "transparent", width = 6, height = 5)
print(Melanoma_dMDT_Library_CDSLength_dMDT)
dev.off()
```


wilcox.test(ENST_Lengths_PepEvidence_dMDT[ENST_Lengths_PepEvidence_dMDT$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_Lengths_PepEvidence_dMDT[ENST_Lengths_PepEvidence_dMDT$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

data:  ENST_Lengths_PepEvidence_dMDT[ENST_Lengths_PepEvidence_dMDT$Evidence == "PeptideEvidence", "Length"]$Length and ENST_Lengths_PepEvidence_dMDT[ENST_Lengths_PepEvidence_dMDT$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 2e+05, p-value = 1e-11
alternative hypothesis: true location shift is not equal to 0

wilcox.test(ENST_CDSLengths_PepEvidence_dMDT[ENST_CDSLengths_PepEvidence_dMDT$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_CDSLengths_PepEvidence_dMDT[ENST_CDSLengths_PepEvidence_dMDT$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

data:  ENST_CDSLengths_PepEvidence_dMDT[ENST_CDSLengths_PepEvidence_dMDT$Evidence == "PeptideEvidence", "Length"]$Length and ENST_CDSLengths_PepEvidence_dMDT[ENST_CDSLengths_PepEvidence_dMDT$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 2e+05, p-value = 2e-10
alternative hypothesis: true location shift is not equal to 0

## cMDT Level Analysis Ovary
```{r IdentifiedvsNonIdentifieddMDTs}
Switches_Ovary_pri <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Pri_GTEx.RDS')
Switches_Ovary_met <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/OvarianCancer/Switches_TuPro_Met_GTEx.RDS')

Switches_Ovary <- rbind(Switches_Ovary_pri, Switches_Ovary_met)
Switches_Ovary_ensps <- merge(ensg_enst_ensp_apris_v104, Switches_Ovary, by.x= 'Transcript.stable.ID', by.y = 'dMDT', sort=FALSE)

ENST_Length_w_Peptides_dMDT_Ov <- ENST_Length_w_Peptides_Ov[ENST_Length_w_Peptides_Ov$ENSP %in% Switches_Ovary_ensps$Protein.stable.ID,]
ENST_Length_wo_Peptides_dMDT_Ov <- ENST_Length_wo_Peptides_Ov[ENST_Length_wo_Peptides_Ov$ENSP %in% Switches_Ovary_ensps$Protein.stable.ID,]

ENST_Lengths_PepEvidence_dMDT_Ov <- rbind(ENST_Length_w_Peptides_dMDT_Ov, ENST_Length_wo_Peptides_dMDT_Ov)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot_dMDT_Ov <- ggplot(ENST_Lengths_PepEvidence_dMDT_Ov , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT_Ov <- ggplot(ENST_Lengths_PepEvidence_dMDT_Ov , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())


Ov_dMDT_Library_TranscriptLength_dMDT <- plot_grid(boxplot_dMDT_Ov, density_plot_dMDT_Ov, align = "v", ncol = 1, rel_heights = c(1, 5))

Ov_dMDT_Library_TranscriptLength_dMDT
svg(filename = "/Users/admin/Desktop/Ov_dMDT_Library_TranscriptLength_dMDT.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Ov_dMDT_Library_TranscriptLength_dMDT.png", plot = Ov_dMDT_Library_TranscriptLength_dMDT, bg = "transparent", width = 6, height = 5)
print(Ov_dMDT_Library_TranscriptLength_dMDT)
dev.off()


## CDS Length 
ENST_CDSLength_w_Peptides_dMDT_Ov <- ENST_CDSLength_w_Peptides_ov[ENST_CDSLength_w_Peptides_ov$ENSP %in% Switches_Ovary_ensps$Protein.stable.ID,]
ENST_CDSLength_wo_Peptides_dMDT_Ov <- ENST_CDSLength_wo_Peptides_ov[ENST_CDSLength_wo_Peptides_ov$ENSP %in% Switches_Ovary_ensps$Protein.stable.ID,]

ENST_CDSLengths_PepEvidence_dMDT_Ov <- rbind(ENST_CDSLength_w_Peptides_dMDT_Ov, ENST_CDSLength_wo_Peptides_dMDT_Ov)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot_dMDT_CDS_ov <- ggplot(ENST_CDSLengths_PepEvidence_dMDT_Ov , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT_CDS_ov <- ggplot(ENST_CDSLengths_PepEvidence_dMDT_Ov , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

library(cowplot)
Ov_dMDT_Library_CDSLength_dMDT_ov <- plot_grid(boxplot_dMDT_CDS_ov, density_plot_dMDT_CDS_ov, align = "v", ncol = 1, rel_heights = c(1, 5))

Ov_dMDT_Library_CDSLength_dMDT_ov


svg(filename = "/Users/admin/Desktop/Ov_dMDT_Library_CDSLength_dMDT_ov.svg", width = 8, height = 6)
ggsave("/Users/admin/Desktop/Ov_dMDT_Library_CDSLength_dMDT_ov.png", plot = Ov_dMDT_Library_CDSLength_dMDT_ov, bg = "transparent", width = 6, height = 5)
print(Ov_dMDT_Library_CDSLength_dMDT_ov)
dev.off()
```


wilcox.test(ENST_Lengths_PepEvidence_dMDT_Ov[ENST_Lengths_PepEvidence_dMDT_Ov$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_Lengths_PepEvidence_dMDT_Ov[ENST_Lengths_PepEvidence_dMDT_Ov$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

data:  ENST_Lengths_PepEvidence_dMDT_Ov[ENST_Lengths_PepEvidence_dMDT_Ov$Evidence == "PeptideEvidence", "Length"]$Length and ENST_Lengths_PepEvidence_dMDT_Ov[ENST_Lengths_PepEvidence_dMDT_Ov$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 1e+05, p-value = 5e-05
alternative hypothesis: true location shift is not equal to 0


wilcox.test(ENST_CDSLengths_PepEvidence_dMDT_Ov[ENST_CDSLengths_PepEvidence_dMDT_Ov$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_CDSLengths_PepEvidence_dMDT_Ov[ENST_CDSLengths_PepEvidence_dMDT_Ov$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

## AML
## cMDT Level Analysis AML
```{r IdentifiedvsNonIdentifieddMDTs}
Switches_aml <- readRDS('/Users/admin/Desktop/TuPro_Paper/Codes/Codes_Figures/Data/cMDT/AML/Switches_TuPro_GTEx.RDS')

Switches_aml_ensps <- merge(ensg_enst_ensp_apris_v104, Switches_aml, by.x= 'Transcript.stable.ID', by.y = 'dMDT', sort=FALSE)

ENST_Length_w_Peptides_dMDT_AML <- ENST_Length_w_Peptides_AML[ENST_Length_w_Peptides_AML$ENSP %in% Switches_aml_ensps$Protein.stable.ID,]

ENST_Length_wo_Peptides_dMDT_AML <- ENST_Length_wo_Peptides_AML[ENST_Length_wo_Peptides_AML$ENSP %in% Switches_aml_ensps$Protein.stable.ID,]

ENST_Lengths_PepEvidence_dMDT_AML <- rbind(ENST_Length_w_Peptides_dMDT_AML, ENST_Length_wo_Peptides_dMDT_AML)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))

density_plot_dMDT_AML <- ggplot(ENST_Lengths_PepEvidence_dMDT_AML , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT_AML <- ggplot(ENST_Lengths_PepEvidence_dMDT_AML , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())


AML_dMDT_Library_TranscriptLength_dMDT <- plot_grid(boxplot_dMDT_AML, density_plot_dMDT_AML, align = "v", ncol = 1, rel_heights = c(1, 5))

AML_dMDT_Library_TranscriptLength_dMDT
svg(filename = "/Users/admin/Desktop/AML_dMDT_Library_TranscriptLength_dMDT.png", width = 8, height = 6)
ggsave("//Users/admin/Desktop/AML_dMDT_Library_TranscriptLength_dMDT.png", plot = AML_dMDT_Library_TranscriptLength_dMDT, bg = "transparent", width = 6, height = 5)
print(AML_dMDT_Library_TranscriptLength_dMDT)
dev.off()


## CDS Length 
ENST_CDSLength_w_Peptides_dMDT_AML <- ENST_CDSLength_w_Peptides_AML[ENST_CDSLength_w_Peptides_AML$ENSP %in% Switches_aml_ensps$Protein.stable.ID,]
ENST_CDSLength_wo_Peptides_dMDT_AML <- ENST_CDSLength_wo_Peptides_AML[ENST_CDSLength_wo_Peptides_AML$ENSP %in% Switches_aml_ensps$Protein.stable.ID,]

ENST_CDSLengths_PepEvidence_dMDT_AML <- rbind(ENST_CDSLength_w_Peptides_dMDT_AML, ENST_CDSLength_wo_Peptides_dMDT_AML)
layout(mat = matrix(c(1,2),2,1, byrow=TRUE),  height = c(1,8))


density_plot_dMDT_CDS_AML <- ggplot(ENST_CDSLengths_PepEvidence_dMDT_AML , aes(x=log10(Length), color=Evidence)) +
  geom_density(lwd = 1.2, linetype = 1) +
  scale_color_manual(values=coul2) + theme_classic()

boxplot_dMDT_CDS_AML <- ggplot(ENST_CDSLengths_PepEvidence_dMDT_AML , aes(y=log10(Length), color=Evidence)) +
  geom_boxplot() + coord_flip() +
  scale_color_manual(values=coul2) + theme_classic() +
  theme(axis.line = element_blank(),
        axis.ticks = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.y = element_blank())

library(cowplot)
aml_dMDT_Library_CDSLength_dMDT_AML <- plot_grid(boxplot_dMDT_CDS_AML, density_plot_dMDT_CDS_AML, align = "v", ncol = 1, rel_heights = c(1, 5))

aml_dMDT_Library_CDSLength_dMDT_AML


svg(filename = "/Users/admin/Desktop/AML_dMDT_Library_CDSLength_dMDT_AML.png", width = 8, height = 6)
ggsave("//Users/admin/Desktop/AML_dMDT_Library_CDSLength_dMDT_AML.png", plot = aml_dMDT_Library_CDSLength_dMDT_AML, bg = "transparent", width = 6, height = 5)
print(aml_dMDT_Library_CDSLength_dMDT_AML)
dev.off()
```

wilcox.test(ENST_Lengths_PepEvidence_dMDT_AML[ENST_Lengths_PepEvidence_dMDT_AML$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_Lengths_PepEvidence_dMDT_AML[ENST_Lengths_PepEvidence_dMDT_AML$Evidence == 'NoPeptideEvidence', 'Length']$Length)
Wilcoxon rank sum test with continuity correction

data:  ENST_Lengths_PepEvidence_dMDT_AML[ENST_Lengths_PepEvidence_dMDT_AML$Evidence == "PeptideEvidence", "Length"]$Length and ENST_Lengths_PepEvidence_dMDT_AML[ENST_Lengths_PepEvidence_dMDT_AML$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 3396, p-value = 0.009
alternative hypothesis: true location shift is not equal to 0

wilcox.test(ENST_CDSLengths_PepEvidence_dMDT_AML[ENST_CDSLengths_PepEvidence_dMDT_AML$Evidence == 'PeptideEvidence', 'Length']$Length, ENST_CDSLengths_PepEvidence_dMDT_AML[ENST_CDSLengths_PepEvidence_dMDT_AML$Evidence == 'NoPeptideEvidence', 'Length']$Length)


Wilcoxon rank sum test with continuity correction

data:  ENST_CDSLengths_PepEvidence_dMDT_AML[ENST_CDSLengths_PepEvidence_dMDT_AML$Evidence == "PeptideEvidence", "Length"]$Length and ENST_CDSLengths_PepEvidence_dMDT_AML[ENST_CDSLengths_PepEvidence_dMDT_AML$Evidence == "NoPeptideEvidence", "Length"]$Length
W = 3074, p-value = 0.01
alternative hypothesis: true location shift is not equal to 0






